---
title: "Heart Disease Or Attack Project - Edx Choose Your Own Submission"
author: "Suhayl Hafiz"
date: "2024-05-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
# set working directory where my data is located, import data.
setwd("//EgnyteDrive/Group/Financial Modelling/Suhayl Tasks/Data Science/Heart Disease Health Indicators/")
```

```{r include=FALSE}
#Import csv from my github directory how please see reference [1] and [2] which is where I downloaded the file from
urlfile="https://raw.githubusercontent.com/suhayl123/EdxCapstone/main/heart_disease_health_indicators_BRFSS2015.csv"
hdhi_df<- read.csv(urlfile)

#Check if we have the required libraries, install them if required and then load them.


if (!require('dplyr')) install.packages('dplyr')
if (!require('Hmisc')) install.packages('Hmisc')
if (!require('caret')) install.packages('caret')
if (!require('scales')) install.packages('scales')
if (!require('ggstats')) install.packages('ggstats')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('fastDummies')) install.packages('fastDummies')
if (!require('cowplot')) install.packages('cowplot')


library(dplyr)
library(Hmisc)
library(caret)
library(scales)
library(ggstats)
library(ggplot2)
library('fastDummies')
library(cowplot)

#To remove scientific notation for number values. 
options(scipen=999)

#Check whether there are any missing values in our dataset i.e. there are none.
hdhi_df %>%summarise_all(funs(sum(is.na(.))))

# To find out the proportion of records where heart disease exits. 9.419% of records has been flagged as having heart disease i.e. there is an imbalance so when it comes to the modelling stage I will create another training dataset that will take all records where heart diesease or attack is present and will randomly under sample records where there is no heart disease to have a more balanced dataset.
summary(hdhi_df$HeartDiseaseorAttack)

#Create histograms of all variables.
hist1 <-hist(hdhi_df$HeartDiseaseorAttack,breaks=n_distinct(hdhi_df$HeartDiseaseorAttack),main ="Histogram of HeartDiseaseorAttack",labels = paste0(round(hist(hdhi_df$HeartDiseaseorAttack, plot = FALSE,breaks = n_distinct(hdhi_df$HeartDiseaseorAttack))$counts / length(hdhi_df$HeartDiseaseorAttack) * 100, 1),"%"))
hist2 <-hist(hdhi_df$HighBP,breaks=n_distinct(hdhi_df$HighBP),main ="Histogram of HighBP",labels = paste0(round(hist(hdhi_df$HighBP, plot = FALSE,breaks = n_distinct(hdhi_df$HighBP))$counts / length(hdhi_df$HighBP) * 100, 1),"%"))
hist3 <-hist(hdhi_df$HighChol,breaks=n_distinct(hdhi_df$HighChol),main ="Histogram of HighChol",labels = paste0(round(hist(hdhi_df$HighChol, plot = FALSE,breaks = n_distinct(hdhi_df$HighChol))$counts / length(hdhi_df$HighChol) * 100, 1),"%"))
hist4 <-hist(hdhi_df$CholCheck,breaks=n_distinct(hdhi_df$CholCheck),main ="Histogram of CholCheck",labels = paste0(round(hist(hdhi_df$CholCheck, plot = FALSE,breaks = n_distinct(hdhi_df$CholCheck))$counts / length(hdhi_df$CholCheck) * 100, 1),"%"))
hist5 <-hist(hdhi_df$Smoker,breaks=n_distinct(hdhi_df$Smoker),main ="Histogram of Smoker",labels = paste0(round(hist(hdhi_df$Smoker, plot = FALSE,breaks = n_distinct(hdhi_df$Smoker))$counts / length(hdhi_df$Smoker) * 100, 1),"%"))
hist6 <-hist(hdhi_df$Stroke,breaks=n_distinct(hdhi_df$Stroke),main ="Histogram of Stroke",labels = paste0(round(hist(hdhi_df$Stroke, plot = FALSE,breaks = n_distinct(hdhi_df$Stroke))$counts / length(hdhi_df$Stroke) * 100, 1),"%"))
hist7 <-hist(hdhi_df$Diabetes,breaks=n_distinct(hdhi_df$Diabetes),main ="Histogram of Diabetes",labels = paste0(round(hist(hdhi_df$Diabetes, plot = FALSE,breaks = n_distinct(hdhi_df$Diabetes))$counts / length(hdhi_df$Diabetes) * 100, 1),"%"))
hist8 <-hist(hdhi_df$PhysActivity,breaks=n_distinct(hdhi_df$PhysActivity),main ="Histogram of PhysActivity",labels = paste0(round(hist(hdhi_df$PhysActivity, plot = FALSE,breaks = n_distinct(hdhi_df$PhysActivity))$counts / length(hdhi_df$PhysActivity) * 100, 1),"%"))
hist9 <-hist(hdhi_df$Fruits,breaks=n_distinct(hdhi_df$Fruits),main ="Histogram of Fruits",labels = paste0(round(hist(hdhi_df$Fruits, plot = FALSE,breaks = n_distinct(hdhi_df$Fruits))$counts / length(hdhi_df$Fruits) * 100, 1),"%"))
hist10 <-hist(hdhi_df$Veggies,breaks=n_distinct(hdhi_df$Veggies),main ="Histogram of Veggies",labels = paste0(round(hist(hdhi_df$Veggies, plot = FALSE,breaks = n_distinct(hdhi_df$Veggies))$counts / length(hdhi_df$Veggies) * 100, 1),"%"))
hist11 <-hist(hdhi_df$HvyAlcoholConsump,breaks=n_distinct(hdhi_df$HvyAlcoholConsump),main ="Histogram of HvyAlcoholConsump",labels = paste0(round(hist(hdhi_df$HvyAlcoholConsump, plot = FALSE,breaks = n_distinct(hdhi_df$HvyAlcoholConsump))$counts / length(hdhi_df$HvyAlcoholConsump) * 100, 1),"%"))
hist12 <-hist(hdhi_df$AnyHealthcare,breaks=n_distinct(hdhi_df$AnyHealthcare),main ="Histogram of AnyHealthcare",labels = paste0(round(hist(hdhi_df$AnyHealthcare, plot = FALSE,breaks = n_distinct(hdhi_df$AnyHealthcare))$counts / length(hdhi_df$AnyHealthcare) * 100, 1),"%"))
hist13 <-hist(hdhi_df$NoDocbcCost,breaks=n_distinct(hdhi_df$NoDocbcCost),main ="Histogram of NoDocbcCost",labels = paste0(round(hist(hdhi_df$NoDocbcCost, plot = FALSE,breaks = n_distinct(hdhi_df$NoDocbcCost))$counts / length(hdhi_df$NoDocbcCost) * 100, 1),"%"))
hist14 <-hist(hdhi_df$GenHlth,breaks=n_distinct(hdhi_df$GenHlth),main ="Histogram of GenHlth",labels = paste0(round(hist(hdhi_df$GenHlth, plot = FALSE,breaks = n_distinct(hdhi_df$GenHlth))$counts / length(hdhi_df$GenHlth) * 100, 1),"%"))
hist15 <-hist(hdhi_df$DiffWalk,breaks=n_distinct(hdhi_df$DiffWalk),main ="Histogram of DiffWalk",labels = paste0(round(hist(hdhi_df$DiffWalk, plot = FALSE,breaks = n_distinct(hdhi_df$DiffWalk))$counts / length(hdhi_df$DiffWalk) * 100, 1),"%"))
hist16 <-hist(hdhi_df$Sex,breaks=n_distinct(hdhi_df$Sex),main ="Histogram of Sex",labels = paste0(round(hist(hdhi_df$Sex, plot = FALSE,breaks = n_distinct(hdhi_df$Sex))$counts / length(hdhi_df$Sex) * 100, 1),"%"))
hist17 <-hist(hdhi_df$BMI,breaks=n_distinct(hdhi_df$BMI),main ="Histogram of BMI",labels = paste0(round(hist(hdhi_df$BMI, plot = FALSE,breaks = n_distinct(hdhi_df$BMI))$counts / length(hdhi_df$BMI) * 100, 1),"%"))
hist18 <-hist(hdhi_df$MentHlth,breaks=n_distinct(hdhi_df$MentHlth),main ="Histogram of MentHlth",labels = paste0(round(hist(hdhi_df$MentHlth, plot = FALSE,breaks = n_distinct(hdhi_df$MentHlth))$counts / length(hdhi_df$MentHlth) * 100, 1),"%"))
hist19 <-hist(hdhi_df$PhysHlth,breaks=n_distinct(hdhi_df$PhysHlth),main ="Histogram of PhysHlth",labels = paste0(round(hist(hdhi_df$PhysHlth, plot = FALSE,breaks = n_distinct(hdhi_df$PhysHlth))$counts / length(hdhi_df$PhysHlth) * 100, 1),"%"))
hist20 <-hist(hdhi_df$Age,breaks=n_distinct(hdhi_df$Age),main ="Histogram of Age",labels = paste0(round(hist(hdhi_df$Age, plot = FALSE,breaks = n_distinct(hdhi_df$Age))$counts / length(hdhi_df$Age) * 100, 1),"%"))
hist21 <-hist(hdhi_df$Income,breaks=n_distinct(hdhi_df$Income),main ="Histogram of Income",labels = paste0(round(hist(hdhi_df$Income, plot = FALSE,breaks = n_distinct(hdhi_df$Income))$counts / length(hdhi_df$Income) * 100, 1),"%"))
hist22 <-hist(hdhi_df$Education,breaks=n_distinct(hdhi_df$Education),main ="Histogram of Education",labels = paste0(round(hist(hdhi_df$Education, plot = FALSE,breaks = n_distinct(hdhi_df$Education))$counts / length(hdhi_df$Education) * 100, 1),"%"))


#Create factor level for HeartDiseaseorAttack.
hdhi_df <- hdhi_df %>% mutate(HDA = case_when(HeartDiseaseorAttack == 0  ~ "NoHDA", HeartDiseaseorAttack == 1  ~ "YesHDA"))
hdhi_df$HDA <- factor(hdhi_df$HDA ,levels =  c("NoHDA","YesHDA"))

#At this current time I will create 3 datasets which are:
#Train - A Training dataset that will be used for model building, this will represent 70% of the entire data
#Test - A test dataset that will be used to assess the perfomance of any model built using the training dataset where I expect to have simillar results in the training dataset, this will represent 15% of the entire data.
#Holdout - A dataset that will not be used for exploratory analysis or model building. This dataset will be used in the final stage to determine how well my final model performed on data that has not been used, this will represent 15% of the entire data
#Create holdout sample.
set.seed(1) 
hold_index <- createDataPartition(y = hdhi_df$HeartDiseaseorAttack, times = 1, p = 0.15, list = FALSE)
Holdout <- hdhi_df[hold_index,]
#Create combined train and test dataset.
TrainTest <- hdhi_df[-hold_index,]
#randomly split the training and test data.
test_index <- createDataPartition(y = TrainTest$HeartDiseaseorAttack, times = 1, p = 0.15/0.85, list = FALSE)
Train <- hdhi_df[-test_index,]
Test <- hdhi_df[test_index,]
#delete indexes and the combined training and test data.
rm( test_index,hold_index,TrainTest)

#Exploratory analysis on training and test data - from the charts it can be see that that someone who has HighBP is more likely to have Heart Disease or Attack.
#HeartDiseaseorAttack by HighBP plots - Train & Test. 
TrainPlot1HighBP<-ggplot(Train) + aes(x = HighBP, fill = factor(HeartDiseaseorAttack), by = factor(HighBP)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By HighBP Group', x = 'HighBP', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2HighBP<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HighBP, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And HighBP', x = 'HighBP', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1HighBP<-ggplot(Test) + aes(x = HighBP, fill = factor(HeartDiseaseorAttack), by = factor(HighBP)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By HighBP Group', x = 'HighBP', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2HighBP<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HighBP, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And HighBP', x = 'HighBP', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by HighChol plots - Train & Test - from the charts it can be see that that someone who has HighChol is more likely to have Heart Disease or Attack.
TrainPlot1HighChol<-ggplot(Train) + aes(x = HighChol, fill = factor(HeartDiseaseorAttack), by = factor(HighChol)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By HighChol Group', x = 'HighChol', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2HighChol<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HighChol, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And HighChol', x = 'HighChol', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1HighChol<-ggplot(Test) + aes(x = HighChol, fill = factor(HeartDiseaseorAttack), by = factor(HighChol)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By HighChol Group', x = 'HighChol', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2HighChol<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HighChol, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And HighChol', x = 'HighChol', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by CholCheck plots - Train & Test - from the charts it can be see that that someone who has CholCheck is more likely to have Heart Disease or Attack however 96% of records have CholCheck with the majority of observations having heart disease or attack falling in this group i.e. likely to not be a significant predictor.
TrainPlot1CholCheck<-ggplot(Train) + aes(x = CholCheck, fill = factor(HeartDiseaseorAttack), by = factor(CholCheck)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By CholCheck Group', x = 'CholCheck', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2CholCheck<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = CholCheck, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And CholCheck', x = 'CholCheck', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1CholCheck<-ggplot(Test) + aes(x = CholCheck, fill = factor(HeartDiseaseorAttack), by = factor(CholCheck)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By CholCheck Group', x = 'CholCheck', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2CholCheck<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = CholCheck, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And CholCheck', x = 'CholCheck', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Smoker plots - Train & Test data. From the charts it can be see that that someone who has smoked is more likely to have Heart Disease or Attack.
TrainPlot1Smoker<-ggplot(Train) + aes(x = Smoker, fill = factor(HeartDiseaseorAttack), by = factor(Smoker)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Smoker Group', x = 'Smoker', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Smoker<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Smoker, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Smoker', x = 'Smoker', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Smoker<-ggplot(Test) + aes(x = Smoker, fill = factor(HeartDiseaseorAttack), by = factor(Smoker)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Smoker Group', x = 'Smoker', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Smoker<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Smoker, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Smoker', x = 'Smoker', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Stroke plots - Train & Test data. From the charts it can be see that that someone who has had a Stroke is significantly likely to have Heart Disease or Attack however this only represent 4% of the data but seems likely that this will be a significant predictor.
TrainPlot1Stroke<-ggplot(Train) + aes(x = Stroke, fill = factor(HeartDiseaseorAttack), by = factor(Stroke)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Stroke Group', x = 'Stroke', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Stroke<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Stroke, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Stroke', x = 'Stroke', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Stroke<-ggplot(Test) + aes(x = Stroke, fill = factor(HeartDiseaseorAttack), by = factor(Stroke)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Stroke Group', x = 'Stroke', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Stroke<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Stroke, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Stroke', x = 'Stroke', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Diabetes plots - Train & TestFrom the charts it can be see that that someone who has diabetes is more likely to have Heart Disease or Attack.
TrainPlot1Diabetes<-ggplot(Train) + aes(x = Diabetes, fill = factor(HeartDiseaseorAttack), by = factor(Diabetes)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Diabetes Group', x = 'Diabetes', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Diabetes<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Diabetes, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Diabetes', x = 'Diabetes', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Diabetes<-ggplot(Test) + aes(x = Diabetes, fill = factor(HeartDiseaseorAttack), by = factor(Diabetes)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Diabetes Group', x = 'Diabetes', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Diabetes<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Diabetes, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Diabetes', x = 'Diabetes', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by PhysActivity plots - Train & Test data. From the charts it can be see that that someone who does Physical Activity on a regular basis is less likely to have Heart Disease or Attack.
TrainPlot1PhysActivity<-ggplot(Train) + aes(x = PhysActivity, fill = factor(HeartDiseaseorAttack), by = factor(PhysActivity)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By PhysActivity Group', x = 'PhysActivity', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2PhysActivity<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = PhysActivity, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And PhysActivity', x = 'PhysActivity', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1PhysActivity<-ggplot(Test) + aes(x = PhysActivity, fill = factor(HeartDiseaseorAttack), by = factor(PhysActivity)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By PhysActivity Group', x = 'PhysActivity', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2PhysActivity<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = PhysActivity, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And PhysActivity', x = 'PhysActivity', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Fruits plots - Train & Test data. From the charts it can be see that that someone who fruit on a daily basis is slightly less likely to have Heart Disease or Attack.
TrainPlot1Fruits<-ggplot(Train) + aes(x = Fruits, fill = factor(HeartDiseaseorAttack), by = factor(Fruits)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Fruits Group', x = 'Fruits', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Fruits<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Fruits, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Fruits', x = 'Fruits', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Fruits<-ggplot(Test) + aes(x = Fruits, fill = factor(HeartDiseaseorAttack), by = factor(Fruits)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Fruits Group', x = 'Fruits', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Fruits<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Fruits, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Fruits', x = 'Fruits', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Veggies plots - Train & Test data. From the charts it can be see that that someone who fruit on a daily basis is less likely to have Heart Disease or Attack.
TrainPlot1Veggies<-ggplot(Train) + aes(x = Veggies, fill = factor(HeartDiseaseorAttack), by = factor(Veggies)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Veggies Group', x = 'Veggies', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Veggies<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Veggies, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Veggies', x = 'Veggies', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Veggies<-ggplot(Test) + aes(x = Veggies, fill = factor(HeartDiseaseorAttack), by = factor(Veggies)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Veggies Group', x = 'Veggies', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Veggies<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Veggies, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Veggies', x = 'Veggies', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by HvyAlcoholConsump plots - Train & Test data. From the charts it can be see that that someone who is a heavy alcohol consumer basis is less likely to have Heart Disease or Attack however only 6% of data is classed as heavy consumer of alcohol.
TrainPlot1HvyAlcoholConsump<-ggplot(Train) + aes(x = HvyAlcoholConsump, fill = factor(HeartDiseaseorAttack), by = factor(HvyAlcoholConsump)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By HvyAlcoholConsump Group', x = 'HvyAlcoholConsump', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2HvyAlcoholConsump<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HvyAlcoholConsump, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And HvyAlcoholConsump', x = 'HvyAlcoholConsump', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1HvyAlcoholConsump<-ggplot(Test) + aes(x = HvyAlcoholConsump, fill = factor(HeartDiseaseorAttack), by = factor(HvyAlcoholConsump)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By HvyAlcoholConsump Group', x = 'HvyAlcoholConsump', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2HvyAlcoholConsump<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = HvyAlcoholConsump, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And HvyAlcoholConsump', x = 'HvyAlcoholConsump', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by AnyHealthcare plots- Train & Test data. From the charts it can be see that that someone who has health care is more likely to have Heart Disease or Attack however 95 % of of records have health care.
TrainPlot1AnyHealthcare<-ggplot(Train) + aes(x = AnyHealthcare, fill = factor(HeartDiseaseorAttack), by = factor(AnyHealthcare)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By AnyHealthcare Group', x = 'AnyHealthcare', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2AnyHealthcare<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = AnyHealthcare, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And AnyHealthcare', x = 'AnyHealthcare', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1AnyHealthcare<-ggplot(Test) + aes(x = AnyHealthcare, fill = factor(HeartDiseaseorAttack), by = factor(AnyHealthcare)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By AnyHealthcare Group', x = 'AnyHealthcare', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2AnyHealthcare<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = AnyHealthcare, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And AnyHealthcare', x = 'AnyHealthcare', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by NoDocbcCost plots - Train & Test data. From the charts it can be see that that someone who could not see a doctor in the last 12 months due to cost is more likely to have Heart Disease or Attack however only 6% of data is classed as heavy consumer of alcohol.
TrainPlot1NoDocbcCost<-ggplot(Train) + aes(x = NoDocbcCost, fill = factor(HeartDiseaseorAttack), by = factor(NoDocbcCost)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By NoDocbcCost Group', x = 'NoDocbcCost', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2NoDocbcCost<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = NoDocbcCost, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And NoDocbcCost', x = 'NoDocbcCost', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1NoDocbcCost<-ggplot(Test) + aes(x = NoDocbcCost, fill = factor(HeartDiseaseorAttack), by = factor(NoDocbcCost)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By NoDocbcCost Group', x = 'NoDocbcCost', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2NoDocbcCost<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = NoDocbcCost, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And NoDocbcCost', x = 'NoDocbcCost', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by GenHlth plots - Train & Test data. From the charts it can be see that that someone who describes them self as having better health is less likely to have Heart Disease or Attack.
TrainPlot1GenHlth<-ggplot(Train) + aes(x = GenHlth, fill = factor(HeartDiseaseorAttack), by = factor(GenHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By GenHlth Group', x = 'GenHlth', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2GenHlth<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = GenHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And GenHlth', x = 'GenHlth', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1GenHlth<-ggplot(Test) + aes(x = GenHlth, fill = factor(HeartDiseaseorAttack), by = factor(GenHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By GenHlth Group', x = 'GenHlth', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2GenHlth<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = GenHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And GenHlth', x = 'GenHlth', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by DiffWalk plots - Train & Test data. From the charts it can be see that that someone who has difficulty walking or climbing stairs is more likely to have Heart Disease or Attack.
TrainPlot1DiffWalk<-ggplot(Train) + aes(x = DiffWalk, fill = factor(HeartDiseaseorAttack), by = factor(DiffWalk)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By DiffWalk Group', x = 'DiffWalk', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2DiffWalk<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = DiffWalk, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And DiffWalk', x = 'DiffWalk', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1DiffWalk<-ggplot(Test) + aes(x = DiffWalk, fill = factor(HeartDiseaseorAttack), by = factor(DiffWalk)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By DiffWalk Group', x = 'DiffWalk', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2DiffWalk<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = DiffWalk, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And DiffWalk', x = 'DiffWalk', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Sex plots - Train & Test data. From the charts it can be see that that males are more likely to have Heart Disease or Attack.
TrainPlot1Sex<-ggplot(Train) + aes(x = Sex, fill = factor(HeartDiseaseorAttack), by = factor(Sex)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Sex Group', x = 'Sex', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Sex<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Sex, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Sex', x = 'Sex', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Sex<-ggplot(Test) + aes(x = Sex, fill = factor(HeartDiseaseorAttack), by = factor(Sex)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Sex Group', x = 'Sex', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Sex<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Sex, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Sex', x = 'Sex', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by BMI plots - Train & Test data. From the charts it we can see that in general the higher a person BMI is the more likely they are to have Heart Disease or Attack howerver this not this does not seem true at the lower end of the BMI scale there I will create a new categorical variable called BMI2 that categorises BMI index into groups which is driven by the NHS (UK).
TrainPlot1BMI<-ggplot(Train) + aes(x = BMI, fill = factor(HeartDiseaseorAttack), by = factor(BMI)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By BMI Group', x = 'BMI', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2BMI<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = BMI, fill = factor(HeartDiseaseorAttack), by = 1)  + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And BMI', x = 'BMI', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1BMI<-ggplot(Test) + aes(x = BMI, fill = factor(HeartDiseaseorAttack), by = factor(BMI)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By BMI Group', x = 'BMI', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2BMI<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = BMI, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And BMI', x = 'BMI', y = 'n obs') + theme(text = element_text(size=7))


#Create BMI2 variable.
Train <- Train %>%
  mutate(BMI2 = case_when(BMI < 18.5  ~ "Underweight",
				   BMI >= 18.5 & BMI <= 24.9  ~ "Healthy", 
				   BMI >= 25 & BMI <= 29.9  ~ "Overweight", 
				   BMI >= 30 & BMI <= 39.9  ~ "Obese", 
				   BMI >= 40   ~ "SeverelyObese", 
				    TRUE ~ "Other"))

Test <- Test %>%
  mutate(BMI2 = case_when(BMI < 18.5  ~ "Underweight",
				   BMI >= 18.5 & BMI <= 24.9  ~ "Healthy", 
				   BMI >= 25 & BMI <= 29.9  ~ "Overweight", 
				   BMI >= 30 & BMI <= 39.9  ~ "Obese", 
				   BMI >= 40   ~ "SeverelyObese", 
				    TRUE ~ "Other"))
#Order the BMI2 categories
Train$BMI2 <- factor(Train$BMI2,levels = c("Underweight", "Healthy", "Overweight", "Obese", "SeverelyObese"))
Test$BMI2 <- factor(Test$BMI2,levels = c("Underweight", "Healthy", "Overweight", "Obese", "SeverelyObese"))

#HeartDiseaseorAttack by BMI2 plots - Train & Test data. From the charts it can be seen clearly that the healthy category is the least likely to have Heart Disease or Attack and that all other categories are more likely. When it comes to modelling I will use the BMI2 variable over BMI.
TrainPlot1BMI2<-ggplot(Train) + aes(x = BMI2, fill = factor(HeartDiseaseorAttack), by = factor(BMI2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By BMI2 Group', x = 'BMI2', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2BMI2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = BMI2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And BMI2', x = 'BMI2', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1BMI2<-ggplot(Test) + aes(x = BMI2, fill = factor(HeartDiseaseorAttack), by = factor(BMI2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By BMI2 Group', x = 'BMI2', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2BMI2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = BMI2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And BMI2', x = 'BMI2', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by MentHlth plots - Train & Test data. The higher the value is (except 0) the more likely you are to have a Heart Disease or Attack however there is there are few observation between 1 and 30 therefore I will create a new variable called MentHlth2 that bins response between 1 and 30 based on visual inspection.
TrainPlot1MentHlth<-ggplot(Train) + aes(x = MentHlth, fill = factor(HeartDiseaseorAttack), by = factor(MentHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By MentHlth Group', x = 'MentHlth', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2MentHlth<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = MentHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And MentHlth', x = 'MentHlth', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1MentHlth<-ggplot(Test) + aes(x = MentHlth, fill = factor(HeartDiseaseorAttack), by = factor(MentHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By MentHlth Group', x = 'MentHlth', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2MentHlth<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = MentHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And MentHlth', x = 'MentHlth', y = 'n obs') + theme(text = element_text(size=7))



#MentHlth2 based on prior plots I've created a new variable:
Train <- Train %>%
  mutate(MentHlth2 = case_when(MentHlth == 0  ~ "0",
				   MentHlth > 0 & MentHlth <= 9  ~ "1to9", 
				   MentHlth > 9 & MentHlth <= 19  ~ "10to19", 
				   MentHlth > 19  ~ "20AndOver", 
				  TRUE ~ "Other"))
Test <- Test %>%
  mutate(MentHlth2 = case_when(MentHlth == 0  ~ "0",
				   MentHlth > 0 & MentHlth <= 9  ~ "1to9", 
				   MentHlth > 9 & MentHlth <= 19  ~ "10to19", 
				   MentHlth > 19  ~ "20AndOver", 
				  TRUE ~ "Other"))
#Order the MentHlth2 categories
Train$MentHlth2 <- factor(Train$MentHlth2,levels = c("0", "1to9", "10to19", "20AndOver"))
Test$MentHlth2 <- factor(Test$MentHlth2,levels = c("0", "1to9", "10to19", "20AndOver"))

#HeartDiseaseorAttack by MentHlth2 plots - Train & Test data. From the charts it can be seen clearly that the 1 to 9 category is the least likely to have Heart Disease or Attack and that all other categories are more likely. When it comes to modelling I will use the MentHlth2 variable over MentHlth.
TrainPlot1MentHlth2<-ggplot(Train) + aes(x = MentHlth2, fill = factor(HeartDiseaseorAttack), by = factor(MentHlth2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By MentHlth2 Group', x = 'MentHlth2', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2MentHlth2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = MentHlth2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And MentHlth2', x = 'MentHlth2', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1MentHlth2<-ggplot(Test) + aes(x = MentHlth2, fill = factor(HeartDiseaseorAttack), by = factor(MentHlth2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By MentHlth2 Group', x = 'MentHlth2', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2MentHlth2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = MentHlth2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And MentHlth2', x = 'MentHlth2', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by PhysHlth plots - Train & Test data. We have a similar trend as MentHlth variable except the trend is more prominent i.e. I will create a new categorical variable called PhysHlth2.
TrainPlot1PhysHlth<-ggplot(Train) + aes(x = PhysHlth, fill = factor(HeartDiseaseorAttack), by = factor(PhysHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By PhysHlth Group', x = 'PhysHlth', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2PhysHlth<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = PhysHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And PhysHlth', x = 'PhysHlth', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1PhysHlth<-ggplot(Test) + aes(x = PhysHlth, fill = factor(HeartDiseaseorAttack), by = factor(PhysHlth)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By PhysHlth Group', x = 'PhysHlth', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2PhysHlth<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = PhysHlth, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And PhysHlth', x = 'PhysHlth', y = 'n obs') + theme(text = element_text(size=7))


#based on plots I've created a new variable:
Train <- Train %>%
  mutate(PhysHlth2 = case_when(PhysHlth >= 0 & PhysHlth <= 2  ~ "0to2",
				   PhysHlth > 2 & PhysHlth <= 9  ~ "3to9", 
				   PhysHlth > 9 & PhysHlth <= 24  ~ "10to24", 
				   PhysHlth > 24 ~ "25AndOver", 
				  TRUE ~ "Other"))


Test <- Test %>%
  mutate(PhysHlth2 = case_when(PhysHlth >= 0 & PhysHlth <= 2  ~ "0to2",
				   PhysHlth > 2 & PhysHlth <= 9  ~ "3to9", 
				   PhysHlth > 9 & PhysHlth <= 24  ~ "10to24", 
				   PhysHlth > 24 ~ "25AndOver", 
				  TRUE ~ "Other"))
#Order the PhysHlth2 categories
Train$PhysHlth2 <- factor(Train$PhysHlth2,levels = c("0to2", "3to9", "10to24", "25AndOver"))
Test$PhysHlth2 <- factor(Test$PhysHlth2,levels = c("0to2", "3to9", "10to24", "25AndOver"))

#HeartDiseaseorAttack by PhysHlth2 plots - Train & Test data. From the charts it can be seen clearly that the lower category the observation is the least likely to have Heart Disease or Attack. When it comes to modelling I will use the PhysHlth2 variable over PhysHlth.
TrainPlot1PhysHlth2<-ggplot(Train) + aes(x = PhysHlth2, fill = factor(HeartDiseaseorAttack), by = factor(PhysHlth2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By PhysHlth2 Group', x = 'PhysHlth2', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2PhysHlth2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = PhysHlth2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And PhysHlth2', x = 'PhysHlth2', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1PhysHlth2<-ggplot(Test) + aes(x = PhysHlth2, fill = factor(HeartDiseaseorAttack), by = factor(PhysHlth2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By PhysHlth2 Group', x = 'PhysHlth2', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2PhysHlth2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = PhysHlth2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And PhysHlth2', x = 'PhysHlth2', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Age plots - Train & Test data. From the charts it can be see that the older a person is the more likely they are to have Heart Disease or Attack.
TrainPlot1Age<-ggplot(Train) + aes(x = Age, fill = factor(HeartDiseaseorAttack), by = factor(Age)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Age Group', x = 'Age', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Age<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Age, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Age', x = 'Age', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Age<-ggplot(Test) + aes(x = Age, fill = factor(HeartDiseaseorAttack), by = factor(Age)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Age Group', x = 'Age', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Age<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Age, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Age', x = 'Age', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Income plots - Train & Test data. From the charts it can be see that the more income a person has the less likely they are to have Heart Disease or Attack.
TrainPlot1Income<-ggplot(Train) + aes(x = Income, fill = factor(HeartDiseaseorAttack), by = factor(Income)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Income Group', x = 'Income', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Income<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Income, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Income', x = 'Income', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Income<-ggplot(Test) + aes(x = Income, fill = factor(HeartDiseaseorAttack), by = factor(Income)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Income Group', x = 'Income', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Income<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Income, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Income', x = 'Income', y = 'n obs') + theme(text = element_text(size=7))

#HeartDiseaseorAttack by Education plots - Train & Test data. From the charts it can be see that the higher the persons education is the less likely they are to have Heart Disease or Attack.
TrainPlot1Education<-ggplot(Train) + aes(x = Education, fill = factor(HeartDiseaseorAttack), by = factor(Education)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Proportion With HDA By Education Group', x = 'Education', y = 'Percent') + theme(text = element_text(size=7))
TrainPlot2Education<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Education, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Education', x = 'Education', y = 'n obs') + theme(text = element_text(size=7))
TestPlot1Education<-ggplot(Test) + aes(x = Education, fill = factor(HeartDiseaseorAttack), by = factor(Education)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Proportion With HDA By Education Group', x = 'Education', y = 'Percent') + theme(text = element_text(size=7))
TestPlot2Education<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot() + aes(x = Education, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Education', x = 'Education', y = 'n obs') + theme(text = element_text(size=7))


#I have decided to create additional variables that combines the outcomes from variables. The first combines HighBP, HighChol, Stroke & Diabetes (1 and 2 will be classed together) to see if we get a different response with relation to Heart Disease or Attack. I feel it's logical to combine these variables as it will measure how severe the persons health is.
#Var 1
Train <- Train %>%
  mutate(Var1 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "BP_Cho", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "BP_Str", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "Cho_Str", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Str_Dia", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "BP_Cho_Str", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Cho_Dia", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Str_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Str_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"))


Test <- Test %>%
  mutate(Var1 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "BP_Cho", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "BP_Str", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "Cho_Str", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Str_Dia", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "BP_Cho_Str", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Cho_Dia", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Str_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Str_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"))

Train$Var1 <- factor(Train$Var1,levels = c("All","BP_Cho_Str","BP_Cho_Dia","BP_Str_Dia","Cho_Str_Dia","Str_Dia","BP_Cho","BP_Str","BP_Dia","Cho_Str","Cho_Dia","Diabetes","HighChol","HighBP","Stroke","None"))
Test$Var1 <- factor(Test$Var1,levels = c("All","BP_Cho_Str","BP_Cho_Dia","BP_Str_Dia","Cho_Str_Dia","Str_Dia","BP_Cho","BP_Str","BP_Dia","Cho_Str","Cho_Dia","Diabetes","HighChol","HighBP","Stroke","None"))

#HeartDiseaseorAttack by Var1 plots - Train & Test data. From the charts it can be see that the more heath conditions a person has the more likely they are to have Heart Disease or Attack except it the person has had a stroke therefore I will create another variable that counts the number of conditions but factors in strokes.
TrainPlot1Var1<-ggplot(Train) + aes(x = Var1, fill = factor(HeartDiseaseorAttack), by = factor(Var1)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Var1 Group', x = 'Var1', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Var1<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var1, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Var1', x = 'Var1', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Var1<-ggplot(Test) + aes(x = Var1, fill = factor(HeartDiseaseorAttack), by = factor(Var1)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Var1 Group', x = 'Var1', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Var1<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var1, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Var1', x = 'Var1', y = 'n obs') + theme(text = element_text(size=7))


#I have created Var2 based on my findings from Var1.
#Var 2
Train <- Train %>%
  mutate(Var2 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All4",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "2ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "3IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "3ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"))


Test <- Test %>%
  mutate(Var2 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All4",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "2ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "3IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "3ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"))

Train$Var2 <- factor(Train$Var2,levels = c("All4","3IncStroke","2IncStroke","Stroke","3ExcStroke","2ExcStroke","Diabetes","HighChol","HighBP","None"))
Test$Var2 <- factor(Test$Var2,levels = c("All4","3IncStroke","2IncStroke","Stroke","3ExcStroke","2ExcStroke","Diabetes","HighChol","HighBP","None"))

#HeartDiseaseorAttack by Var2 plots - Train & Test data. From the charts it can be proven that my findings from Var1 is justified (the more health conditions has a person has the more likely they are to have Heart Disease or Attack) except if the person has had a stroke only vs 3 conditions excluding a stroke. 
TrainPlot1Var2<-ggplot(Train) + aes(x = Var2, fill = factor(HeartDiseaseorAttack), by = factor(Var2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Var2 Group', x = 'Var2', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Var2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Var2', x = 'Var2', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Var2<-ggplot(Test) + aes(x = Var2, fill = factor(HeartDiseaseorAttack), by = factor(Var2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Var2 Group', x = 'Var2', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Var2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Var2', x = 'Var2', y = 'n obs') + theme(text = element_text(size=7))




#The 3rd variable I have created combines Veggies and Fruits variables together.
#Var3
Train <- Train %>%
  mutate(Var3 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "Fruits", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "None", 
				    TRUE ~ "Other"))

Test <- Test %>%
  mutate(Var3 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "Fruits", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "None", 
				    TRUE ~ "Other"))

Train$Var3 <- factor(Train$Var3,levels = c("Fru_Veg","Fruits","Veggies","None"))
Test$Var3 <- factor(Test$Var3,levels = c("Fru_Veg","Fruits","Veggies","None"))

#HeartDiseaseorAttack by Var3 plots - Train & Test data. From the charts you can see that if a person has Fruits and Veggies everyday they are the least likely to have Heart Disease or Attack then by Veggies only, I have decided to group Fruits only and None together as they have a similar response to Heart Disease or Attack.
TrainPlot1Var3<-ggplot(Train) + aes(x = Var3, fill = factor(HeartDiseaseorAttack), by = factor(Var3)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Var3 Group', x = 'Var3', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Var3<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var3, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Var3', x = 'Var3', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Var3<-ggplot(Test) + aes(x = Var3, fill = factor(HeartDiseaseorAttack), by = factor(Var3)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Var3 Group', x = 'Var3', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Var3<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var3, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Var3', x = 'Var3', y = 'n obs') + theme(text = element_text(size=7))

#The 4th variable I have created is based on my learnings from Var3.
#Var4
Train <- Train %>%
  mutate(Var4 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "NoVegies", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "NoVegies", 
				    TRUE ~ "Other"))

Test <- Test %>%
  mutate(Var4 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "NoVegies", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "NoVegies", 
				    TRUE ~ "Other"))

Train$Var4 <- factor(Train$Var4,levels = c("Fru_Veg","Veggies","NoVegies"))
Test$Var4 <- factor(Test$Var4,levels =  c("Fru_Veg","Veggies","NoVegies"))

#HeartDiseaseorAttack by Var4 plots - Train & Test data. From the charts you will see that it agrees with my findings from Var3. This will be used for modelling rather than Var3, Fruits & Veggies.
TrainPlot1Var4<-ggplot(Train) + aes(x = Var4, fill = factor(HeartDiseaseorAttack), by = factor(Var4)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Var4 Group', x = 'Var4', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Var4<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var4, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Var4', x = 'Var4', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Var4<-ggplot(Test) + aes(x = Var4, fill = factor(HeartDiseaseorAttack), by = factor(Var4)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Var4 Group', x = 'Var4', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Var4<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var4, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Var4', x = 'Var4', y = 'n obs') + theme(text = element_text(size=7))


#The 5th variable I have created combines HvyAlcoholConsump and Smoker variables together.
Train <- Train %>%
  mutate(Var5 = case_when(HvyAlcoholConsump == 1 & Smoker ==1 ~ "HAC_Smo",
				    HvyAlcoholConsump == 1 & Smoker ==0 ~ "HvyAlcoholConsump", 
				    HvyAlcoholConsump == 0 & Smoker ==1 ~ "Smoker",  
				    HvyAlcoholConsump == 0 & Smoker ==0 ~ "None", 
				    TRUE ~ "Other"))

Test <- Test %>%
  mutate(Var5 = case_when(HvyAlcoholConsump == 1 & Smoker ==1 ~ "HAC_Smo",
				    HvyAlcoholConsump == 1 & Smoker ==0 ~ "HvyAlcoholConsump", 
				    HvyAlcoholConsump == 0 & Smoker ==1 ~ "Smoker",  
				    HvyAlcoholConsump == 0 & Smoker ==0 ~ "None", 
				    TRUE ~ "Other"))

Train$Var5 <- factor(Train$Var5,levels = c("HAC_Smo","HvyAlcoholConsump","None","Smoker"))
Test$Var5 <- factor(Test$Var5,levels =  c("HAC_Smo","HvyAlcoholConsump","None","Smoker"))

#HeartDiseaseorAttack by Var5 plots - Train & Test data. From the charts I can see that only smoking produces the highest risk to Heart Disease or Attack. Suprisingly heavy alcohol use produces the lowest however this could just mean that person suffers from other diseases and doesn't get a chance to get Heart Disease or Attack.
TrainPlot1Var5<-ggplot(Train) + aes(x = Var5, fill = factor(HeartDiseaseorAttack), by = factor(Var5)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Var5 Group', x = 'Var5', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Var5<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var5, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Var5', x = 'Var5', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Var5<-ggplot(Test) + aes(x = Var5, fill = factor(HeartDiseaseorAttack), by = factor(Var5)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Var5 Group', x = 'Var5', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Var5<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Var5, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Var5', x = 'Var5', y = 'n obs') + theme(text = element_text(size=7))




#The 6th variable I have created combines education groups 1 and 3 together since there small proportion of these people in these groups and seem to be more at risk
Train <- Train %>%
  mutate(Education2 = case_when(Education >=1  & Education<=3 ~ "1_3", 
				Education ==4  ~ "4",
				Education ==5  ~ "5",
				Education ==6  ~ "6",
				TRUE ~ "Other"))

Test <- Test %>%
  mutate(Education2 = case_when(Education >=1  & Education<=3 ~ "1_3", 
				Education ==4  ~ "4",
				Education ==5  ~ "5",
				Education ==6  ~ "6",
				TRUE ~ "Other"))

Train$Education2 <- factor(Train$Education2,levels = c("1_3","4","5","6"))
Test$Education2 <- factor(Test$Education2,levels =  c("1_3","4","5","6"))

#HeartDiseaseorAttack by Education2 plots - Train & Test data. From the charts I can see that only smoking produces the highest risk to Heart Disease or Attack. Suprisingly heavy alcohol use produces the lowest however this could just mean that person suffers from other diseases and doesn't get a chance to get Heart Disease or Attack.
TrainPlot1Education2<-ggplot(Train) + aes(x = Education2, fill = factor(HeartDiseaseorAttack), by = factor(Education2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Education2 Group', x = 'Education2', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Education2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Education2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Education2', x = 'Education2', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Education2<-ggplot(Test) + aes(x = Education2, fill = factor(HeartDiseaseorAttack), by = factor(Education2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Education2 Group', x = 'Education2', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Education2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Education2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Education2', x = 'Education2', y = 'n obs') + theme(text = element_text(size=7))


#The 7th variable I have created combines Age groups 1 to 4, 5 to 6, 7, 8 to 9, 10, 11 to 12 and 13 
Train <- Train %>%
  mutate(Age2 = case_when(Age >=1  & Age<=4 ~ "1_4", 
				Age >=5  & Age<=6 ~ "5_6", 
				Age ==7  ~ "7", 
				Age >=8  & Age<=9 ~ "8_9", 
				Age ==10  ~ "10", 
				Age >=11  & Age<=12 ~ "11_12", 
				Age ==13 ~ "13", 
				TRUE ~ "Other"))

Test <- Test %>%
  mutate(Age2 = case_when(Age >=1  & Age<=4 ~ "1_4", 
				Age >=5  & Age<=6 ~ "5_6", 
				Age ==7  ~ "7", 
				Age >=8  & Age<=9 ~ "8_9", 
				Age ==10  ~ "10", 
				Age >=11  & Age<=12 ~ "11_12", 
				Age ==13 ~ "13", 
				TRUE ~ "Other"))

Train$Age2 <- factor(Train$Age2 ,levels = c("1_4","5_6","7","8_9","10","11_12","13"))
Test$Age2 <- factor(Test$Age2 ,levels =  c("1_4","5_6","7","8_9","10","11_12","13"))

#HeartDiseaseorAttack by Age2 plots - Train & Test data. From the charts I can see that only smoking produces the highest risk to Heart Disease or Attack. Suprisingly heavy alcohol use produces the lowest however this could just mean that person suffers from other diseases and doesn't get a chance to get Heart Disease or Attack.
TrainPlot1Age2<-ggplot(Train) + aes(x = Age2, fill = factor(HeartDiseaseorAttack), by = factor(Age2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Train - Proportion With HDA By Age2 Group', x = 'Age2', y = 'Percent')  + theme(text = element_text(size=7))
TrainPlot2Age2<-Train %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Age2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Train - Total Proportion With HDA And Age2', x = 'Age2', y = 'n obs')  + theme(text = element_text(size=7))
TestPlot1Age2<-ggplot(Test) + aes(x = Age2, fill = factor(HeartDiseaseorAttack), by = factor(Age2)) + geom_bar(position = "fill") + geom_text(stat = "prop", position = position_fill(.5))  + labs(title = 'Test - Proportion With HDA By Age2 Group', x = 'Age2', y = 'Percent')   + theme(text = element_text(size=7))
TestPlot2Age2<-Test %>% filter(HeartDiseaseorAttack==1) %>% ggplot()  + aes(x = Age2, fill = factor(HeartDiseaseorAttack), by = 1) + geom_bar() +geom_text(aes(label = scales::percent(after_stat(prop), accuracy = 1)),stat = "prop", position = position_stack(.5)) + labs(fill='HeartDiseaseorAttack',  title = 'Test - Total Proportion With HDA And Age2', x = 'Age2', y = 'n obs') + theme(text = element_text(size=7))


#convert to factors
names <- c('Diabetes' ,'GenHlth','Income')
Train[,names] <- lapply(Train[,names] , factor)
Test[,names] <- lapply(Test[,names] , factor)

```

```{r include=FALSE}
#Since we have a few observations where we have heart disease attack as occurring i.e. I have imbalanced data with regards to Heart Disease or Attack. I can balance the Heart Disease or Attack outcome by either adding in random samples of observations where heart disease is present or I can remove observations where heart disease isn't present. I have decided to under sample and will keep 20% of where records that do not contain HeartDiseaseorAttack for the fact that I will need less computing power to run any modelling. Once any model has been created I will assess the model against my test data and the original training data.

TrainKeep1 <- Train %>% filter(HeartDiseaseorAttack==1)
TrainKeep2 <- Train %>% filter(HeartDiseaseorAttack==0)
#Create a random index that will keep only 20% of data held in TrainKeep2 (HeartDiseaseorAttack = 0)
Train_index <- createDataPartition(y = TrainKeep2$BMI, times = 1, p = 0.2, list = FALSE)
set.seed(1)
TrainKeep2 <- TrainKeep2[Train_index,]
#Now combine TrainKeep1 & TrainKeep2 as Train2
Train2 <- rbind(TrainKeep1, TrainKeep2)
rm(Train_index,TrainKeep1, TrainKeep2)
#you can see that I now have a 34.13% response rate in my dataset that I will be using for model building rather than 9.39% i.e. the imbalance has been reduced.
summary(Train2$HeartDiseaseorAttack)


#I'm now going to create 4 models using the caret package (Random Forest, Logistic Regression and CART). Based on the best model I will then look to improve that model.
#cv set to 3 - computer power
trainControl <- trainControl(method="cv", number=3,summaryFunction = twoClassSummary,classProbs = TRUE)


# RF
set.seed(1)
fit.rf <- train(HDA~ HighBP+ HighChol+ CholCheck+ Smoker+ Stroke+  Diabetes+ PhysActivity+ HvyAlcoholConsump+ AnyHealthcare+ NoDocbcCost+ GenHlth+ DiffWalk+ Sex+ Age2+ Education2 + Income  + Var1 + Var2 + Var3 + Var4 + Var5+ BMI2  + MentHlth2 + PhysHlth2 , data=Train2, method = "rf", metric="ROC",trControl = trainControl,importance = TRUE)
#model summary results Train
y_hat_fit.rf_Train <- predict(fit.rf,Train,type="raw")
cm.rf_Train <- confusionMatrix(y_hat_fit.rf_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.rf_Test <- predict(fit.rf,Test,type="raw")
cm.rf_Test <- confusionMatrix(y_hat_fit.rf_Test, factor(Test$HDA))
cm.rf_Train
cm.rf_Test
#sommers d 
Train.rf.pr <- as.data.frame(y_hat_fit.rf_Train)
Train.rf.pr <- Train.rf.pr  %>% mutate(y = case_when(y_hat_fit.rf_Train  == 'YesHDA' ~ 1, y_hat_fit.rf_Train  == 'NoHDA'~ 0 ))
Test.rf.pr <- as.data.frame(y_hat_fit.rf_Test)
Test.rf.pr <- Test.rf.pr  %>% mutate(y = case_when(y_hat_fit.rf_Test  == 'YesHDA'  ~ 1, y_hat_fit.rf_Test == 'NoHDA'~ 0 ))
somers2(Train.rf.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.rf.pr$y,  Test$HeartDiseaseorAttack)


# LG - Logistic Regression
set.seed(1)
fit.glm <- train(HDA~ HighBP+ HighChol+ CholCheck+ Smoker+ Stroke+  Diabetes+ PhysActivity+ HvyAlcoholConsump+ AnyHealthcare+ NoDocbcCost+ GenHlth+ DiffWalk+ Sex+ Age2+ Education2 + Income  + Var1 + Var2 + Var3 + Var4 + Var5+ BMI2  + MentHlth2 + PhysHlth2 , data=Train2 , method="glmStepAIC",family="binomial", direction="forward", metric="ROC", trControl=trainControl)
#model summary results Train
y_hat_fit.glm_Train <- predict(fit.glm,Train,type="raw")
cm.glm_Train <- confusionMatrix(y_hat_fit.glm_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm_Test <- predict(fit.glm,Test,type="raw")
cm.glm_Test <- confusionMatrix(y_hat_fit.glm_Test, factor(Test$HDA))
cm.glm_Train
cm.glm_Test
#sommers d 
Train.glm.pr <- as.data.frame(y_hat_fit.glm_Train)
Train.glm.pr <- Train.glm.pr  %>% mutate(y = case_when(y_hat_fit.glm_Train  == 'YesHDA' ~ 1, y_hat_fit.glm_Train  == 'NoHDA'~ 0 ))
Test.glm.pr <- as.data.frame(y_hat_fit.glm_Test)
Test.glm.pr <- Test.glm.pr  %>% mutate(y = case_when(y_hat_fit.glm_Test  == 'YesHDA' ~ 1, y_hat_fit.glm_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm.pr$y,  Test$HeartDiseaseorAttack)


# CART - Classification and Regression Trees (CART), 
set.seed(1)
fit.cart <- train(HDA~ HighBP+ HighChol+ CholCheck+ Smoker+ Stroke+  Diabetes+ PhysActivity+ HvyAlcoholConsump+ AnyHealthcare+ NoDocbcCost+ GenHlth+ DiffWalk+ Sex+ Age2+ Education2 + Income  + Var1 + Var2 + Var3 + Var4 + Var5+ BMI2  + MentHlth2 + PhysHlth2, data=Train2, method="rpart", metric="ROC",trControl=trainControl)
#model summary results Train
y_hat_fit.cart_Train <- predict(fit.cart,Train,type="raw")
cm.cart_Train <- confusionMatrix(y_hat_fit.cart_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.cart_Test <- predict(fit.cart,Test,type="raw")
cm.cart_Test <- confusionMatrix(y_hat_fit.cart_Test, factor(Test$HDA))
cm.cart_Train
cm.cart_Test
#sommers d 
Train.cart.pr <- as.data.frame(y_hat_fit.cart_Train)
Train.cart.pr <- Train.cart.pr  %>% mutate(y = case_when(y_hat_fit.cart_Train  == 'YesHDA' ~ 1, y_hat_fit.cart_Train  == 'NoHDA'~ 0 ))
Test.cart.pr <- as.data.frame(y_hat_fit.cart_Test)
Test.cart.pr <- Test.cart.pr  %>% mutate(y = case_when(y_hat_fit.cart_Test  == 'YesHDA' ~ 1, y_hat_fit.cart_Test  == 'NoHDA'~ 0 ))
somers2(Train.cart.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.cart.pr$y,  Test$HeartDiseaseorAttack)


#create a dataframe to compare Nvars, Accuracy,Sensitivity (note Sensitivity and Specificity are incorrectly populated in the table), Precision & Somers D

ModelResultsSummary <- data.frame(
	Model = c("RF","GLM","CART"),  
	NVars= c(length(fit.rf$finalModel$importance[,1]),length(row(varImp(fit.glm$finalModel)[1])),length(fit.cart$finalModel$variable.importance)),
	TrainAccuracy = c(cm.rf_Train$overall["Accuracy"],cm.glm_Train$overall["Accuracy"],cm.cart_Train$overall["Accuracy"]),
	TestAccuracy = c(cm.rf_Test$overall["Accuracy"],cm.glm_Test$overall["Accuracy"],cm.cart_Test$overall["Accuracy"]),
	TrainSensitivity = c(cm.rf_Train$byClass["Specificity"],cm.glm_Train$byClass["Specificity"],cm.cart_Train$byClass["Specificity"]),
	TestSensitivity = c(cm.rf_Test$byClass["Specificity"],cm.glm_Test$byClass["Specificity"],cm.cart_Test$byClass["Specificity"]),
	TrainPrecision = c(cm.rf_Train$byClass["Neg Pred Value"],cm.glm_Train$byClass["Neg Pred Value"],cm.cart_Train$byClass["Neg Pred Value"]),
	TestPrecision = c(cm.rf_Test$byClass["Neg Pred Value"],cm.glm_Test$byClass["Neg Pred Value"],cm.cart_Test$byClass["Neg Pred Value"]),
	TrainNcountTP = c(cm.rf_Train$table[4],cm.glm_Train$table[4],cm.cart_Train$table[4]),
	TestNcountTP = c(cm.rf_Test$table[4],cm.glm_Test$table[4],cm.cart_Test$table[4]),
	TrainSomersD = c(somers2(Train.rf.pr$y,  Train$HeartDiseaseorAttack)["Dxy"],somers2(Train.glm.pr$y,  Train$HeartDiseaseorAttack)["Dxy"],somers2(Train.cart.pr$y,  Train$HeartDiseaseorAttack)["Dxy"]),
	TestSomersD = c(somers2(Test.rf.pr$y,  Test$HeartDiseaseorAttack)["Dxy"],somers2(Test.glm.pr$y,  Test$HeartDiseaseorAttack)["Dxy"],somers2(Test.cart.pr$y,  Test$HeartDiseaseorAttack)["Dxy"])
)
#I have chosen to explore glm model further.
```

```{r include=FALSE}
#I have chosen to explore glm model further.

#create dummy variables & train GLM using the all the variables that are present in my first model using forward selection.
Train <- dummy_cols(Train, select_columns = c('Var1','Var2', 'Var3','Var4','Var5','Diabetes','GenHlth','Education2','Income','BMI2','Age2','MentHlth2','PhysHlth2'), remove_selected_columns = FALSE)
Train2 <- dummy_cols(Train2, select_columns = c('Var1','Var2', 'Var3','Var4','Var5','Diabetes','GenHlth','Education2','Income','BMI2','Age2','MentHlth2','PhysHlth2'), remove_selected_columns = FALSE)
Test <- dummy_cols(Test, select_columns = c('Var1','Var2', 'Var3','Var4','Var5','Diabetes','GenHlth','Education2','Income','BMI2','Age2','MentHlth2','PhysHlth2'), remove_selected_columns = FALSE)

#increase cv to 7
trainControl <- trainControl(method="cv", number=7,summaryFunction = twoClassSummary,classProbs = TRUE)

# Now I will retrain but using the dummy variables where possbile based on the variables in the 1st GLM model
fit.glm2 <- train(HDA~ Var1_All+ Var1_BP_Cho_Str+ Var1_BP_Cho_Dia+ Var1_BP_Str_Dia+ Var1_Cho_Str_Dia+ Var1_Str_Dia+ Var1_BP_Cho+ Var1_BP_Str+ Var1_BP_Dia+ Var1_Cho_Str+ Var1_Cho_Dia+  Var1_Diabetes+ 
Var1_HighChol+ Var1_HighBP+ Var1_Stroke+  Var1_None+ DiffWalk+ Sex+ Stroke+ Age2_1_4+ Age2_5_6+  Age2_7+  Age2_8_9+  Age2_10+  Age2_11_12+ Age2_13+ GenHlth_1+ GenHlth_2+ 
GenHlth_3+ GenHlth_4+ GenHlth_5+ HighChol+ Var5_HAC_Smo+  Var5_HvyAlcoholConsump+  Var5_None+  Var5_Smoker+ Income_1+Income_2+Income_3+Income_4+ Income_5+Income_6+
Income_7+Income_8 + HighBP + NoDocbcCost+ CholCheck + Education2_1_3 + Education2_4 + Education2_5 + Education2_6 + PhysHlth2_0to2+ PhysHlth2_3to9+ PhysHlth2_10to24+ PhysHlth2_25AndOver +  BMI2_Healthy+  BMI2_Overweight+  BMI2_Obese+ 
BMI2_SeverelyObese+ MentHlth2_0+ MentHlth2_1to9+ MentHlth2_10to19+ MentHlth2_20AndOver + Var3_Fru_Veg+  Var3_Fruits+ Var3_None+  Var3_Veggies+ AnyHealthcare+ PhysActivity + Diabetes_0+ Diabetes_1+ Diabetes_2 ,
data=Train2 , method="glmStepAIC",family="binomial", direction="forward", metric="ROC", trControl=trainControl)
#model summary results Train
y_hat_fit.glm2_Train <- predict(fit.glm2,Train,type="raw")
cm.glm2_Train <- confusionMatrix(y_hat_fit.glm2_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm2_Test <- predict(fit.glm2,Test,type="raw")
cm.glm2_Test <- confusionMatrix(y_hat_fit.glm2_Test, factor(Test$HDA))
cm.glm2_Train
cm.glm2_Test
#sommers d 
Train.glm2.pr <- as.data.frame(y_hat_fit.glm2_Train)
Train.glm2.pr <- Train.glm2.pr  %>% mutate(y = case_when(y_hat_fit.glm2_Train  == 'YesHDA' ~ 1, y_hat_fit.glm2_Train  == 'NoHDA'~ 0 ))
Test.glm2.pr <- as.data.frame(y_hat_fit.glm2_Test)
Test.glm2.pr <- Test.glm2.pr  %>% mutate(y = case_when(y_hat_fit.glm2_Test  == 'YesHDA' ~ 1, y_hat_fit.glm2_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm2.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm2.pr$y,  Test$HeartDiseaseorAttack)
summary(fit.glm2)

#similar results produce, I will now create another model using the same variables except I will now stepwiseAIC procedure to determine which variables are most important in order and will use the dummy variables as opppsed to any categorical variable.

ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM2", NVars=length(row(varImp(fit.glm2$finalModel)[1])),  TrainAccuracy = cm.glm2_Train$overall["Accuracy"], TestAccuracy =cm.glm2_Test$overall["Accuracy"] ,TrainSensitivity = cm.glm2_Train$byClass["Specificity"], TestSensitivity = cm.glm2_Test$byClass["Specificity"], TrainPrecision = cm.glm2_Train$byClass["Neg Pred Value"], TestPrecision = cm.glm2_Test$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm2_Train$table[4] , TestNcountTP = cm.glm2_Test$table[4],TrainSomersD = somers2(Train.glm2.pr$y,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test.glm2.pr$y,  Test$HeartDiseaseorAttack)["Dxy"] )

#Run backwards with the variable selected from the forward selection
fit.glm3 <- train(HDA~ Var1_None + DiffWalk + Sex + Age2_1_4 + Stroke + Age2_5_6 + GenHlth_1 + GenHlth_2 + GenHlth_3 + Age2_13 + Age2_11_12 + Var5_Smoker + HighChol + Var1_HighChol + Age2_10 + GenHlth_4 + Age2_7 + Var1_HighBP + Income_8 + HighBP + NoDocbcCost + CholCheck + PhysHlth2_10to24 +
 Diabetes_2 + Income_2 + Var1_Stroke + Income_1 + PhysHlth2_0to2 + Education2_5 + Income_3 + Var1_Str_Dia + Var1_BP_Str_Dia + BMI2_Healthy + Income_4 + Var3_Fruits + Var1_All + AnyHealthcare + Var1_Cho_Dia + MentHlth2_20AndOver + PhysActivity + Var5_HvyAlcoholConsump,
data=Train2 , method="glmStepAIC",family="binomial", direction="backward", metric="ROC", trControl=trainControl)

#model summary results Train
y_hat_fit.glm3_Train <- predict(fit.glm3,Train,type="raw")
cm.glm3_Train <- confusionMatrix(y_hat_fit.glm3_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm3_Test <- predict(fit.glm3,Test,type="raw")
cm.glm3_Test <- confusionMatrix(y_hat_fit.glm3_Test, factor(Test$HDA))
cm.glm3_Train
cm.glm3_Test
#sommers d 
Train.glm3.pr <- as.data.frame(y_hat_fit.glm3_Train)
Train.glm3.pr <- Train.glm3.pr  %>% mutate(y = case_when(y_hat_fit.glm3_Train  == 'YesHDA' ~ 1, y_hat_fit.glm3_Train  == 'NoHDA'~ 0 ))
Test.glm3.pr <- as.data.frame(y_hat_fit.glm3_Test)
Test.glm3.pr <- Test.glm3.pr  %>% mutate(y = case_when(y_hat_fit.glm3_Test  == 'YesHDA' ~ 1, y_hat_fit.glm3_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm3.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm3.pr$y,  Test$HeartDiseaseorAttack)
#No Change compared to prior glm model using forward selection
summary(fit.glm3)


ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM3", NVars=length(row(varImp(fit.glm3$finalModel)[1])),  TrainAccuracy = cm.glm3_Train$overall["Accuracy"], TestAccuracy =cm.glm3_Test$overall["Accuracy"] ,TrainSensitivity = cm.glm3_Train$byClass["Specificity"], TestSensitivity = cm.glm3_Test$byClass["Specificity"], TrainPrecision = cm.glm3_Train$byClass["Neg Pred Value"], TestPrecision = cm.glm3_Test$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm3_Train$table[4] , TestNcountTP = cm.glm3_Test$table[4],TrainSomersD = somers2(Train.glm3.pr$y,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test.glm3.pr$y,  Test$HeartDiseaseorAttack)["Dxy"] )

#StepAIC model
fit.glm4 <- train(HDA~ Var1_All+ Var1_BP_Cho_Str+ Var1_BP_Cho_Dia+ Var1_BP_Str_Dia+ Var1_Cho_Str_Dia+ Var1_Str_Dia+ Var1_BP_Cho+ Var1_BP_Str+ Var1_BP_Dia+ Var1_Cho_Str+ Var1_Cho_Dia+  Var1_Diabetes+ 
Var1_HighChol+ Var1_HighBP+ Var1_Stroke+  Var1_None+ DiffWalk+ Sex+ Stroke+ Age2_1_4+ Age2_5_6+  Age2_7+  Age2_8_9+  Age2_10+  Age2_11_12+ Age2_13+ GenHlth_1+ GenHlth_2+ 
GenHlth_3+ GenHlth_4+ GenHlth_5+ HighChol+ Var5_HAC_Smo+  Var5_HvyAlcoholConsump+  Var5_None+  Var5_Smoker+ Income_1+Income_2+Income_3+Income_4+ Income_5+Income_6+
Income_7+Income_8 + HighBP + NoDocbcCost+ CholCheck + Education2_1_3 + Education2_4 + Education2_5 + Education2_6 + PhysHlth2_0to2+ PhysHlth2_3to9+ PhysHlth2_10to24+ PhysHlth2_25AndOver +  BMI2_Healthy+  BMI2_Overweight+  BMI2_Obese+ 
BMI2_SeverelyObese+ MentHlth2_0+ MentHlth2_1to9+ MentHlth2_10to19+ MentHlth2_20AndOver + Var3_Fru_Veg+  Var3_Fruits+ Var3_None+  Var3_Veggies+ AnyHealthcare+ PhysActivity + Diabetes_0+ Diabetes_1+ Diabetes_2 ,
data=Train2 , method="glmStepAIC",family="binomial", direction="both", metric="ROC", trControl=trainControl)

#model summary results Train
y_hat_fit.glm4_Train <- predict(fit.glm4,Train,type="raw")
cm.glm4_Train <- confusionMatrix(y_hat_fit.glm4_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm4_Test <- predict(fit.glm4,Test,type="raw")
cm.glm4_Test <- confusionMatrix(y_hat_fit.glm4_Test, factor(Test$HDA))
cm.glm4_Train
cm.glm4_Test
#sommers d 
Train.glm4.pr <- as.data.frame(y_hat_fit.glm4_Train)
Train.glm4.pr <- Train.glm4.pr  %>% mutate(y = case_when(y_hat_fit.glm4_Train  == 'YesHDA' ~ 1, y_hat_fit.glm4_Train  == 'NoHDA'~ 0 ))
Test.glm4.pr <- as.data.frame(y_hat_fit.glm4_Test)
Test.glm4.pr <- Test.glm4.pr  %>% mutate(y = case_when(y_hat_fit.glm4_Test  == 'YesHDA' ~ 1, y_hat_fit.glm4_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm4.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm4.pr$y,  Test$HeartDiseaseorAttack)
summary(fit.glm4)

ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM4", NVars=length(row(varImp(fit.glm4$finalModel)[1])),  TrainAccuracy = cm.glm4_Train$overall["Accuracy"], TestAccuracy =cm.glm4_Test$overall["Accuracy"] ,TrainSensitivity = cm.glm4_Train$byClass["Specificity"], TestSensitivity = cm.glm4_Test$byClass["Specificity"], TrainPrecision = cm.glm4_Train$byClass["Neg Pred Value"], TestPrecision = cm.glm4_Test$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm4_Train$table[4] , TestNcountTP = cm.glm4_Test$table[4],TrainSomersD = somers2(Train.glm4.pr$y,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test.glm4.pr$y,  Test$HeartDiseaseorAttack)["Dxy"] )
summary(fit.glm4$finalModel)


#The forward and backward models produced the best results however there are insigniifcant variables i.e. I will create a 5th model removing insignificant variables from model glm3.fit

fit.glm5 <- train(HDA~ Var1_None + DiffWalk + Sex + Age2_1_4 + Stroke + Age2_5_6 + GenHlth_1 + GenHlth_2 + GenHlth_3 + Age2_13 + Age2_11_12 + Var5_Smoker + HighChol + Var1_HighChol + Age2_10 + GenHlth_4 + Age2_7 + Var1_HighBP + Income_8 + HighBP + NoDocbcCost + CholCheck + PhysHlth2_10to24 + Diabetes_2 + Income_2 + Var1_Stroke + Income_1 + PhysHlth2_0to2 + Education2_5 + Income_3 + Var1_Str_Dia + Var1_BP_Str_Dia + BMI2_Healthy + Income_4 + Var3_Fruits + Var1_All ,
data=Train2 , method="glm",family="binomial", metric="ROC", trControl=trainControl)

#model summary results Train
y_hat_fit.glm5_Train <- predict(fit.glm5,Train,type="raw")
cm.glm5_Train <- confusionMatrix(y_hat_fit.glm5_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm5_Test <- predict(fit.glm5,Test,type="raw")
cm.glm5_Test <- confusionMatrix(y_hat_fit.glm5_Test, factor(Test$HDA))
cm.glm5_Train
cm.glm5_Test
#sommers d 
Train.glm5.pr <- as.data.frame(y_hat_fit.glm5_Train)
Train.glm5.pr <- Train.glm5.pr  %>% mutate(y = case_when(y_hat_fit.glm5_Train  == 'YesHDA' ~ 1, y_hat_fit.glm5_Train  == 'NoHDA'~ 0 ))
Test.glm5.pr <- as.data.frame(y_hat_fit.glm5_Test)
Test.glm5.pr <- Test.glm5.pr  %>% mutate(y = case_when(y_hat_fit.glm5_Test  == 'YesHDA' ~ 1, y_hat_fit.glm5_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm5.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm5.pr$y,  Test$HeartDiseaseorAttack)
summary(fit.glm5)

ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM5", NVars=length(row(varImp(fit.glm5$finalModel)[1])),  TrainAccuracy = cm.glm5_Train$overall["Accuracy"], TestAccuracy =cm.glm5_Test$overall["Accuracy"] ,TrainSensitivity = cm.glm5_Train$byClass["Specificity"], TestSensitivity = cm.glm5_Test$byClass["Specificity"], TrainPrecision = cm.glm5_Train$byClass["Neg Pred Value"], TestPrecision = cm.glm5_Test$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm5_Train$table[4] , TestNcountTP = cm.glm5_Test$table[4],TrainSomersD = somers2(Train.glm5.pr$y,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test.glm5.pr$y,  Test$HeartDiseaseorAttack)["Dxy"] )
summary(fit.glm5$finalModel)

# Create a 6th model removing insignificant variables from model glm4.fit
#DO the same for Stepwise (model 4) 
fit.glm6 <- train(HDA~ Var1_All + Var1_BP_Cho_Str + Var1_BP_Cho_Dia + Var1_BP_Str_Dia + Var1_Cho_Str_Dia + Var1_Str_Dia + Var1_BP_Cho + Var1_BP_Str + Var1_BP_Dia + Var1_Cho_Str + Var1_Cho_Dia + Var1_Diabetes + Var1_HighChol + Var1_HighBP + Var1_Stroke + DiffWalk + Sex + Age2_1_4 + Age2_5_6 + Age2_7 + Age2_8_9 + Age2_10 + Age2_11_12 + GenHlth_1 + GenHlth_2 + GenHlth_3 + GenHlth_4 + Var5_HAC_Smo + Var5_HvyAlcoholConsump + Var5_None + Income_1 + Income_2 + Income_3 + Income_4 + Income_5 + Income_6 + Income_7 + NoDocbcCost + CholCheck + Education2_5 + PhysHlth2_0to2 + PhysHlth2_10to24 + BMI2_Healthy + Var3_Fruits + Diabetes_1,
                  data=Train2 , method="glm",family="binomial", metric="ROC", trControl=trainControl)

#model summary results Train
y_hat_fit.glm6_Train <- predict(fit.glm6,Train,type="raw")
cm.glm6_Train <- confusionMatrix(y_hat_fit.glm6_Train, factor(Train$HDA))
#model summary results Test
y_hat_fit.glm6_Test <- predict(fit.glm6,Test,type="raw")
cm.glm6_Test <- confusionMatrix(y_hat_fit.glm6_Test, factor(Test$HDA))
cm.glm6_Train
cm.glm6_Test
#sommers d 
Train.glm6.pr <- as.data.frame(y_hat_fit.glm6_Train)
Train.glm6.pr <- Train.glm6.pr  %>% mutate(y = case_when(y_hat_fit.glm6_Train  == 'YesHDA' ~ 1, y_hat_fit.glm6_Train  == 'NoHDA'~ 0 ))
Test.glm6.pr <- as.data.frame(y_hat_fit.glm6_Test)
Test.glm6.pr <- Test.glm6.pr  %>% mutate(y = case_when(y_hat_fit.glm6_Test  == 'YesHDA' ~ 1, y_hat_fit.glm6_Test  == 'NoHDA'~ 0  ))
somers2(Train.glm6.pr$y,  Train$HeartDiseaseorAttack)
somers2(Test.glm6.pr$y,  Test$HeartDiseaseorAttack)
summary(fit.glm6)

#Update model results summary 
ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM6", NVars=length(row(varImp(fit.glm6$finalModel)[1])),  TrainAccuracy = cm.glm6_Train$overall["Accuracy"], TestAccuracy =cm.glm6_Test$overall["Accuracy"] ,TrainSensitivity = cm.glm6_Train$byClass["Specificity"], TestSensitivity = cm.glm6_Test$byClass["Specificity"], TrainPrecision = cm.glm6_Train$byClass["Neg Pred Value"], TestPrecision = cm.glm6_Test$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm6_Train$table[4] , TestNcountTP = cm.glm6_Test$table[4],TrainSomersD = somers2(Train.glm6.pr$y,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test.glm6.pr$y,  Test$HeartDiseaseorAttack)["Dxy"] )


#I have decided that my final model will be the 5th GLM model.
summary(fit.glm5$finalModel)

#I will now discover what is my best probability score should be to determine the highest somers d statistic in my train and test data.
#Train
fit.glmF_P_Train <- predict(fit.glm5,Train,type="prob")
names(fit.glmF_P_Train)[1] <- "PscoreHDA0"
names(fit.glmF_P_Train)[2] <- "PscoreHDA1"
fit.glmF_P_Train<- data.frame(Train$HeartDiseaseorAttack, fit.glmF_P_Train$PscoreHDA1)
names(fit.glmF_P_Train)[1] <- "HeartDiseaseorAttack"
names(fit.glmF_P_Train)[2] <- "PscoreHDA1"
#Test
fit.glmF_P_Test <- predict(fit.glm5,Test,type="prob")
names(fit.glmF_P_Test)[1] <- "PscoreHDA0"
names(fit.glmF_P_Test)[2] <- "PscoreHDA1"
fit.glmF_P_Test<- data.frame(Test$HeartDiseaseorAttack, fit.glmF_P_Test$PscoreHDA1)
names(fit.glmF_P_Test)[1] <- "HeartDiseaseorAttack"
names(fit.glmF_P_Test)[2] <- "PscoreHDA1"


#Find best cutoff using Train
alphas <- seq(0.00, 1, 0.01)
cutoffsTrain <- sapply(alphas, function(alpha){
	fit.glmF_P_Train <- fit.glmF_P_Train %>% mutate(PredRes = case_when (PscoreHDA1 < alpha ~ 0, PscoreHDA1 >= alpha ~ 1)) 
	somers2(fit.glmF_P_Train$PredRes,  Train$HeartDiseaseorAttack)
})
cutoffsTrain <- as.data.frame(t(cutoffsTrain[,-1]))

#Same for Test
cutoffsTest <- sapply(alphas, function(alpha){
	fit.glmF_P_Test <- fit.glmF_P_Test %>% mutate(PredRes = case_when (PscoreHDA1 < alpha ~ 0, PscoreHDA1 >= alpha ~ 1)) 
	somers2(fit.glmF_P_Test$PredRes,  Test$HeartDiseaseorAttack)
})
cutoffsTest <- as.data.frame(t(cutoffsTest[,-1]))

# My cutoff values for my training and test are very similar
par(mfrow = c(1, 2))
plot(cutoffsTrain$Dxy)
plot(cutoffsTest$Dxy)
cutoffsTrain[which.max(cutoffsTrain$Dxy),]
cutoffsTest[which.max(cutoffsTest$Dxy),]

#I've decided to take the average best cut off values from my train & test result.
cutoffvalue <- (0.30 + 0.33) / 2 

# assess how well my cut off value performs using my new cut off value 0.31
y_hat_p_fit.glm5_Train <- predict(fit.glm5,Train,type="prob")
names(y_hat_p_fit.glm5_Train)[1] <- "PscoreHDA0GLM5"
names(y_hat_p_fit.glm5_Train)[2] <- "PscoreHDA1GLM5"
#Test
y_hat_p_fit.glm5_Test <- predict(fit.glm5,Test,type="prob")
names(y_hat_p_fit.glm5_Test)[1] <- "PscoreHDA0GLM5"
names(y_hat_p_fit.glm5_Test)[2] <- "PscoreHDA1GLM5"

#Train Summary
Train_Results <- Train[1]
Train_Results <- bind_cols(Train_Results, y_hat_p_fit.glm5_Train)
#based on my highest cuttoff value in my training 
Train_Results <- Train_Results %>% mutate(ResultHDA1GLM5 = case_when (PscoreHDA1GLM5 < cutoffvalue~ 0, PscoreHDA1GLM5 >= cutoffvalue~ 1))
#For graphical view
Train_Results <- Train_Results %>% mutate(PredCatGLM5 = case_when (PscoreHDA1GLM5 <= 0.1 ~ 1, PscoreHDA1GLM5 <= 0.2 ~ 2, PscoreHDA1GLM5 <= 0.3 ~ 3, PscoreHDA1GLM5 <= 0.4 ~ 4, PscoreHDA1GLM5 <= 0.5 ~ 5, PscoreHDA1GLM5 <= 0.6 ~ 6, PscoreHDA1GLM5 <= 0.7 ~ 7, PscoreHDA1GLM5 <= 0.8 ~ 8, PscoreHDA1GLM5 <= 0.9 ~ 9, PscoreHDA1GLM5 > 0.9 ~ 10))
GLM5SummaryResultsTrain<-Train_Results %>%group_by(PredCatGLM5) %>%summarise(Freq = n(),PredProb = mean(PscoreHDA1GLM5),Target = mean(HeartDiseaseorAttack))


#Test Summary
Test_Results <- Test[1]
Test_Results <- bind_cols(Test_Results, y_hat_p_fit.glm5_Test)
#based on my highest cuttoff value in my training 
Test_Results <- Test_Results %>% mutate(ResultHDA1GLM5 = case_when (PscoreHDA1GLM5 < cutoffvalue ~ 0, PscoreHDA1GLM5 >= cutoffvalue~ 1))
#For graphical view
Test_Results <- Test_Results %>% mutate(PredCatGLM5 = case_when (PscoreHDA1GLM5 <= 0.1 ~ 1, PscoreHDA1GLM5 <= 0.2 ~ 2, PscoreHDA1GLM5 <= 0.3 ~ 3, PscoreHDA1GLM5 <= 0.4 ~ 4, PscoreHDA1GLM5 <= 0.5 ~ 5, PscoreHDA1GLM5 <= 0.6 ~ 6, PscoreHDA1GLM5 <= 0.7 ~ 7, PscoreHDA1GLM5 <= 0.8 ~ 8, PscoreHDA1GLM5 <= 0.9 ~ 9, PscoreHDA1GLM5 > 0.9 ~ 10))
GLM5SummaryResultsTest<-Test_Results %>%group_by(PredCatGLM5) %>%summarise(Freq = n(),PredProb = mean(PscoreHDA1GLM5),Target = mean(HeartDiseaseorAttack))

#based on my cutoff value here is my confusion matrix 

#model summary results Train
cm.glm5_TrainOptimised <- confusionMatrix(factor(Train_Results$ResultHDA1GLM5), factor(Train_Results$HeartDiseaseorAttack))
#model summary results Test
cm.glm5_TestOptimised <- confusionMatrix(factor(Test_Results$ResultHDA1GLM5), factor(Test_Results$HeartDiseaseorAttack))
cm.glm5_TrainOptimised
cm.glm5_TestOptimised
#sommers d 
somers2(Train_Results$ResultHDA1GLM5,  Train$HeartDiseaseorAttack)
somers2(Test_Results$ResultHDA1GLM5,  Test$HeartDiseaseorAttack)

#Update model summary results, since I'm using HeartDiseaseorAttack in my confusion matrix rather than HDA variable Sensitivity and Precision are not marked the wrong way round.
ModelResultsSummary <- ModelResultsSummary %>% add_row(Model = "GLM5 Optimised", NVars=length(row(varImp(fit.glm5$finalModel)[1])),  TrainAccuracy = cm.glm5_TrainOptimised$overall["Accuracy"], TestAccuracy =cm.glm5_TestOptimised$overall["Accuracy"] ,TrainSensitivity = cm.glm5_TrainOptimised$byClass["Specificity"], TestSensitivity = cm.glm5_TestOptimised$byClass["Specificity"], TrainPrecision = cm.glm5_TrainOptimised$byClass["Neg Pred Value"], TestPrecision = cm.glm5_TestOptimised$byClass["Neg Pred Value"] , TrainNcountTP = cm.glm5_TrainOptimised$table[4] , TestNcountTP = cm.glm5_TestOptimised$table[4],TrainSomersD = somers2(Train_Results$ResultHDA1GLM5,  Train$HeartDiseaseorAttack)["Dxy"], TestSomersD =somers2(Test_Results$ResultHDA1GLM5,  Test$HeartDiseaseorAttack)["Dxy"] )

#Since I have decided to go with with my 5th GLM model that is optimised with regards to Somers D, I will now see how the model performs on the holdout data.
```

```{r include=FALSE}

#I will now discover what is my best probability score should be to determine the highest somers d statistic in my train and test data.
#Train
fit.glmF_P_Train <- predict(fit.glm5,Train,type="prob")
names(fit.glmF_P_Train)[1] <- "PscoreHDA0"
names(fit.glmF_P_Train)[2] <- "PscoreHDA1"
fit.glmF_P_Train<- data.frame(Train$HeartDiseaseorAttack, fit.glmF_P_Train$PscoreHDA1)
names(fit.glmF_P_Train)[1] <- "HeartDiseaseorAttack"
names(fit.glmF_P_Train)[2] <- "PscoreHDA1"
#Test
fit.glmF_P_Test <- predict(fit.glm5,Test,type="prob")
names(fit.glmF_P_Test)[1] <- "PscoreHDA0"
names(fit.glmF_P_Test)[2] <- "PscoreHDA1"
fit.glmF_P_Test<- data.frame(Test$HeartDiseaseorAttack, fit.glmF_P_Test$PscoreHDA1)
names(fit.glmF_P_Test)[1] <- "HeartDiseaseorAttack"
names(fit.glmF_P_Test)[2] <- "PscoreHDA1"


#Prep my holdout data by adding all the variables I created in the Train and Test datasets.
Holdout <- Holdout %>%
  mutate(BMI2 = case_when(BMI < 18.5  ~ "Underweight",
				   BMI >= 18.5 & BMI <= 24.9  ~ "Healthy", 
				   BMI >= 25 & BMI <= 29.9  ~ "Overweight", 
				   BMI >= 30 & BMI <= 39.9  ~ "Obese", 
				   BMI >= 40   ~ "SeverelyObese", 
				    TRUE ~ "Other"),
	MentHlth2 = case_when(MentHlth == 0  ~ "0",
				   MentHlth > 0 & MentHlth <= 9  ~ "1to9", 
				   MentHlth > 9 & MentHlth <= 19  ~ "10to19", 
				   MentHlth > 19  ~ "20AndOver", 
				  TRUE ~ "Other"),
	Var1 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "BP_Cho", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "BP_Str", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "Cho_Str", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Str_Dia", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "BP_Cho_Str", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Cho_Dia", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "BP_Str_Dia", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "Cho_Str_Dia", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"),
	Var2 = case_when(HighBP == 1 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "All4",
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "HighBP", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "HighChol", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "Stroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "Diabetes", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & Diabetes ==0 ~ "2ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "2IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "2ExcStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "2IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 1 & Diabetes ==0 ~ "3IncStroke", 
				    HighBP == 1 & HighChol ==1 & Stroke == 0 & (Diabetes ==1 | Diabetes ==2) ~ "3ExcStroke", 
				    HighBP == 1 & HighChol ==0 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==1 & Stroke == 1 & (Diabetes ==1 | Diabetes ==2) ~ "3IncStroke", 
				    HighBP == 0 & HighChol ==0 & Stroke == 0 & Diabetes ==0 ~ "None", 
				    TRUE ~ "Other"),
	Var3 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "Fruits", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "None", 
				    TRUE ~ "Other"),
	Var4 = case_when(Fruits == 1 & Veggies ==1 ~ "Fru_Veg",
				    Fruits == 1 & Veggies ==0 ~ "NoVegies", 
				    Fruits == 0 & Veggies ==1 ~ "Veggies", 
				    Fruits == 0 & Veggies ==0 ~ "NoVegies", 
				    TRUE ~ "Other"),
	Var5 = case_when(HvyAlcoholConsump == 1 & Smoker ==1 ~ "HAC_Smo",
				    HvyAlcoholConsump == 1 & Smoker ==0 ~ "HvyAlcoholConsump", 
				    HvyAlcoholConsump == 0 & Smoker ==1 ~ "Smoker",  
				    HvyAlcoholConsump == 0 & Smoker ==0 ~ "None", 
				    TRUE ~ "Other"),
	Education2 = case_when(Education >=1  & Education<=3 ~ "1_3", 
				Education ==4  ~ "4",
				Education ==5  ~ "5",
				Education ==6  ~ "6",
				TRUE ~ "Other"),
	Age2 = case_when(Age >=1  & Age<=4 ~ "1_4", 
				Age >=5  & Age<=6 ~ "5_6", 
				Age ==7  ~ "7", 
				Age >=8  & Age<=9 ~ "8_9", 
				Age ==10  ~ "10", 
				Age >=11  & Age<=12 ~ "11_12", 
				Age ==13 ~ "13", 
				TRUE ~ "Other")	,
	PhysHlth2 = case_when(PhysHlth >= 0 & PhysHlth <= 2  ~ "0to2",
				   PhysHlth > 2 & PhysHlth <= 9  ~ "3to9", 
				   PhysHlth > 9 & PhysHlth <= 24  ~ "10to24", 
				   PhysHlth > 24 ~ "25AndOver" ,
				   TRUE ~ "Other")
)

#Create factor levels for the holdout
Holdout$BMI2 <- factor(Holdout$BMI2,levels = c("Underweight", "Healthy", "Overweight", "Obese", "SeverelyObese"))
Holdout$MentHlth2 <- factor(Holdout$MentHlth2,levels = c("0", "1to9", "10to19", "20AndOver"))
Holdout$PhysHlth2 <- factor(Holdout$PhysHlth2,levels = c("0to2", "3to9", "10to24", "25AndOver"))
Holdout$Var1 <- factor(Holdout$Var1,levels = c("All","BP_Cho_Str","BP_Cho_Dia","BP_Str_Dia","Cho_Str_Dia","Str_Dia","BP_Cho","BP_Str","BP_Dia","Cho_Str","Cho_Dia","Diabetes","HighChol","HighBP","Stroke","None"))
Holdout$Var2 <- factor(Holdout$Var2,levels = c("All4","3IncStroke","2IncStroke","Stroke","3ExcStroke","2ExcStroke","Diabetes","HighChol","HighBP","None"))
Holdout$Var3 <- factor(Holdout$Var3,levels = c("Fru_Veg","Fruits","Veggies","None"))
Holdout$Var4 <- factor(Holdout$Var4,levels = c("Fru_Veg","Veggies","NoVegies"))
Holdout$Var5 <- factor(Holdout$Var5,levels = c("HAC_Smo","HvyAlcoholConsump","None","Smoker"))
Holdout$Education2 <- factor(Holdout$Education2,levels = c("1_3","4","5","6"))
Holdout$Age2 <- factor(Holdout$Age2 ,levels = c("1_4","5_6","7","8_9","10","11_12","13"))

#create dummy vars for my holdout
Holdout <- dummy_cols(Holdout, select_columns =c('Var1','Var2', 'Var3','Var4','Var5','Diabetes','GenHlth','Education2','Income','BMI2','Age2','MentHlth2','PhysHlth2'), remove_selected_columns = FALSE)


#using the holdout data get use fit.GLM5 model to predict the probability of whether observations will have heart disease or attack.
y_hat_p_fit.glm5_Holdout <- predict(fit.glm5,Holdout,type="prob")
names(y_hat_p_fit.glm5_Holdout)[1] <- "PscoreHDA0GLM5"
names(y_hat_p_fit.glm5_Holdout)[2] <- "PscoreHDA1GLM5"


#create Holdout Summary
Holdout_Results <- Holdout[1]
Holdout_Results <- bind_cols(Holdout_Results, y_hat_p_fit.glm5_Holdout)
#based on my cuttoff value 
Holdout_Results <- Holdout_Results %>% mutate(ResultHDA1GLM5 = case_when (PscoreHDA1GLM5 < cutoffvalue ~ 0, PscoreHDA1GLM5 >= cutoffvalue ~ 1))

# create 10 bins for predicted probability score
Holdout_Results <- Holdout_Results %>% mutate(PredCatGLM5 = case_when (PscoreHDA1GLM5 <= 0.1 ~ 1, PscoreHDA1GLM5 <= 0.2 ~ 2, PscoreHDA1GLM5 <= 0.3 ~ 3, PscoreHDA1GLM5 <= 0.4 ~ 4, PscoreHDA1GLM5 <= 0.5 ~ 5, PscoreHDA1GLM5 <= 0.6 ~ 6, PscoreHDA1GLM5 <= 0.7 ~ 7, PscoreHDA1GLM5 <= 0.8 ~ 8, PscoreHDA1GLM5 <= 0.9 ~ 9, PscoreHDA1GLM5 > 0.9 ~ 10))
GLM5SummaryResultsHoldout<-Holdout_Results %>%group_by(PredCatGLM5) %>%summarise(Freq = n(),PredProb = mean(PscoreHDA1GLM5),Target = mean(HeartDiseaseorAttack))

#model summary results Holdout
cm.glm5_HoldoutOptimised <- confusionMatrix(factor(Holdout_Results$ResultHDA1GLM5), factor(Holdout_Results$HeartDiseaseorAttack))
cm.glm5_HoldoutOptimised

#sommers d 
somers2(Holdout_Results$ResultHDA1GLM5,  Holdout$HeartDiseaseorAttack)
```

# Introduction

The following paragraph has been taken from the Kaggle website and has been reduced [1] (the link can be found in my reference section) to give you some context, Heart Disease is among the most prevalent chronic diseases in the United States which impacts millions of Americans each year and exerts significant financial burden on the economy. There are different types of coronary heart disease, the majority of individuals only learn they have the disease following symptoms such as chest pain, a heart attack, or sudden cardiac arrest. The Behavioral Risk Factor Surveillance System (BRFSS) is a health-related telephone survey that is collected annually by the CDC. Each year, the survey collects responses from over 400,000 Americans on health-related risk behaviors, chronic health conditions, and the use of preventative services. It has been conducted every year since 1984.

For this project, I've downloaded a csv dataset available on Kaggle for the year 2015 which has been cleaned. The dataset contains 253,680 survey responses and contains the following variables:

-   HeartDiseaseorAttack: This is my target variable that I will be predicting Respondents that have ever reported having coronary heart disease (CHD) or myocardial infarction (MI). This is binary variable which contains 23,893 people that have had heart disease or attack. Therefore we have an overall response rate of 9.419% i.e. we have an imbalance and will need to take this into account carrying out modelling.

-   HighBP - High Blood Pressure:Adults who have been told they have high blood pressure by a doctor, nurse, or other health professional. O means no and 1 means yes

-   HighChol - High Cholesterol: Have you EVER been told by a doctor, nurse or other health professional that your blood cholesterol is high? O means no and 1 means yes.

-   CholCheck: checked cholesterol in past 5 years by a doctor, nurse or other health professional that your blood cholesterol is high. O means no and 1 means yes.

-   BMI - BMI: BMI to nearest percentage.

-   Smoking - Smoker: Whether the person smoked atleast 100 cigarettes in your entire life? [Note: 5 packs = 100 cigarettes]. 0 means no and 1 means yes.

-   Stroke: (Ever told) you had a stroke. O means no and 1 means yes.

-   Diabetes - (Ever told) you have diabetes. 0 is for no diabetes or only during pregnancy, 1 is for pre-diabetes or borderline diabetes, 2 is for yes diabetes.

-   PhysActivity - Physical Activity: Adults who reported doing physical activity or exercise during the past 30 days other than their regular job. 1 for physical activity, 0 for no physical activity.

-   Fruits: 0 means no fruit consumed per day. 1 means consumed 1 or more pieces of fruit per day.

-   Veggies: 0.this means no vegetables consumed per day. 1 will mean consumed 1 or more pieces of vegetable per day.

-   HvyAlcoholConsump - Alcohol Consumption: Heavy drinkers (adult men having more than 14 drinks per week and adult women having more than 7 drinks per week). 0 means not heavy drinker and 1 means heavy drinker.

-   AnyHealthcare: Do you have any kind of health care coverage, including health insurance, prepaid plans such as HMOs, or government plans such as Medicare, or Indian Health Service? . 0 means no and 1 means yes.

-   NoDocbcCost: Was there a time in the past 12 months when you needed to see a doctor but could not because of cost? . 0 means no and 1 means yes.

-   GenHlth: Health General and Mental Health: - Would you say that in general your health is: --\> GENHLTH. This is made of 5 categories, 1 is Excellent , 2 is Very Good, 3 is good, 4 is fair and 5 is poor.

-   MentHlth: Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good? in days from 0-30.

-   PhysHlth: Now thinking about your physical health, which includes physical illness and injury, for how many days during the past 30 days was your physical health not good? in days from 0-30.

-   DiffWalk - Do you have serious difficulty walking or climbing stairs? --\> DIFFWALK . 0 means no and 1 means yes.

-   Sex: sex of respondent. female is 0 and male is 1.

-   Age: age category - 1 is 18-24 all the way up to 13 wis 80 and older. 5 year increments.

-   Education: What is the highest grade or year of school you completed? --\> EDUCA. 1 being never attended school or kindergarten only up to 6 being college 4 years or more .

-   Income: Is your annual household income from all sources: (If respondent refuses at any income level, code "Refused.") --\> INCOME2. 1 being less than \$10,000 all the way up to 8 being \$75,000 or more

Here are histograms of all the variables listed above:

```{r echo=FALSE}
par(mfrow = c(3, 3))
plot(hist1)
plot(hist2)
plot(hist3)
plot(hist4)
plot(hist5)
plot(hist6)
plot(hist7)
plot(hist8)
plot(hist9)
par(mfrow = c(3, 3))
plot(hist10)
plot(hist11)
plot(hist12)
plot(hist13)
plot(hist14)
plot(hist15)
plot(hist16)
par(mfrow = c(2, 3))
plot(hist17)
plot(hist18)
plot(hist19)
plot(hist20)
plot(hist21)
plot(hist22)
```

The aim of this project is to create a predictive model that identifies which respondents have heart disease based and determines which people are most likely to have heart disease or attack based on the variables I described above.

# Methods/Analysis

I decided to create 4 datasets which come from the original dataframe (hdhi_df) and each observation does not crossover into another dataset (except between Train & Train2).

1)  Train - This represents 70% of the original dataframe and has been randomly selected. Going forward this dataset will be used for any exploratory analysis and to assess any models that have been built from my Train2 dataset.

2)  Train2 - Since I have imbalance data with regards to Heart Disease or Attack, I can fix this by either adding in random samples of observations where heart disease is present or I can remove observations where heart disease isn't present. I have decided to under sample since this requires less computational requirements for model building. This dataset comes from my Train dataset only and it contains all observations where Heart Disease or Attack is present and I have randomly selected 20% of observations where Heart Disease or Attack is not present.

3)  Test - This represents 15% of the original dataframe and has been randomly selected. Going forward this dataset will be used to compare any exploratory analysis and model building against my trainning set.

4)  Holdout - This remaining 15% of the original dataframe goes intom my holdout dataset which will be used only at the final stage to determine how well my final model performed i.e. the records will not be used in any model building.

Here is a summary table of all the 4 datasets which shows you the number of records and the average response rate of observations with heart diesease or attack:

```{r echo=FALSE}
data.frame(dataset = c("Train","Train2","Test","Holdout"), N_Records=c(nrow(Train), nrow(Train2), nrow(Test), nrow(Holdout)), MeanHDA = c(mean(Train$HeartDiseaseorAttack), mean(Train2$HeartDiseaseorAttack), mean(Test$HeartDiseaseorAttack) , mean(Holdout$HeartDiseaseorAttack)))
```

### Exploratory Analysis

Before any model building was done I did exploratory analysis to first of all to understand my data more, to see if could create my own variables based on the variables I have and if I could simplify any variables by reducing the number of categories (all variables are categorical except BMI). I created a 2 by 2 graphical output of all variables to compare my Train and Test datasets which shows the average response rate of heart disease or attack by each value held in the variable that I'm exploring followed by the proportion of records held in each value by heart disease or attack.Here is an example below using the Education variable:

```{r echo=FALSE}
plots<-plot_grid(TrainPlot1Education + theme(legend.position = "none"),  TestPlot1Education + theme(legend.position = "none") + ylab(""), TrainPlot2Education + theme(legend.position = "none") , TestPlot2Education + theme(legend.position = "none") +  ylab(""))
legend <- get_legend(TrainPlot1Education)
plot_grid(plots, legend, rel_widths = c(2, 0.2))
```

As you can see from the above chart I can reduce the number of categories in the Education variable if I group 1 and 3 levels together since their response rates to heart disease or attack are similar in the Train and Test datasets and kept groups 4, 5 & 6 separately i.e. I created a new variable called Education2 which will replace Education in the model building step.

Based on the variables that I've been given (see introduction) I created these additional variables using the training and test datasets.

-   Var1 - I have created this variable which is the outcome of the responses from HighBP, HighChol, Stroke and Diabetes variables.

-   Var2 - I have created this variable based on the information I gathered from Var1, which basically counts the no. of conditions a person has from HighBP, HighChol, Stroke and Diabetes but includes and excludes people with a stroke.

-   Var3 - I have created this variable based on outcomes from Fruits and Veggies variables.This variable replaces the Fruits and Veggies variables when it came to the model building stage.

-   Var4 - I have created this variable based on the information I gathered from Var3, which groups people who eat fruits and veggies consistently, have veggies and do not have veggies consistently. This variable replaces the Fruits and Veggies variables when it came to the model building stage.

-   Var5 - I have created this variable which is the outcome of the responses from HvyAlcoholConsump, Smoker variables.

-   BMI2- I have created this variable which is the outcome of the BMI variable, I have converted the numeric responses into bands which I got from the NHS [3]. Below 18.5 -- underweight, 18.5 to 24.9 -- healthy weight, 25 to 29.9 -- overweight, 30 to 39.9 -- obese and 40 or above -- severely obese. This variable replaces the BMI variable when it came to the model building stage.

-   MentHlth2 - I have created this variable which is based on the outcome MentHlth with the following categories 0, 1 to 9, 10 to 19 & 20 and over. This variable replaces the MentHlth variable when it came to the model building stage.

-   PhysHlth2 - I have created this variable which is based on the outcome PhysHlth2 with the following categories 0 to 2, 3 to 9, 10 to 24 & 25 and over. This variable replaces the PhysHlth variable when it came to the model building stage.

-   Education2 - I have created this variable which is based on the outcome from Education with the following categories 1 to 3, 4, 5 and 6. This variable replaces the Education variable when it came to the model building stage.

-   Age2- I have created this variable which is based on the outcome from Age with the following categories 1 to 4, 5 to 6, 7, 8 to 9, 10, 11 to 12 and 13. This variable replaces the Age variable when it came to the model building stage.

I have provided you charts below to show you the response rates for by category for each of the categorical responses using my Train dataset only:

```{r echo=FALSE}
plot_grid(TrainPlot1Var1 + theme(legend.position="none") , TrainPlot1Var2 + theme(legend.position="none"), ncol = 1, align = 'v') 
```

As you can see from the above charts when a person has had a stroke they more likely to suffer from heart disease or attack which you can see in the 1st chart above. Therefire it makes sense to create Var2 based on Var1 since the observations where a stoke is present they are more likely to suffer from heart disease or attack irregardless of how many conditions they have from diabetes, high cholesterol or high blood pressure.

```{r echo=FALSE}
plots<-plot_grid(TrainPlot1Var3 + theme(legend.position = "none"),  TrainPlot1Var4 + theme(legend.position = "none") + ylab(""),  TrainPlot1Var5 + theme(legend.position = "none") , TrainPlot1BMI2 + theme(legend.position = "none") +  ylab(""))
legend <- get_legend(TrainPlot1Education)
plot_grid(plots, legend, rel_widths = c(2, 0.2))
```

Similarly with Var3 I saw that I could reduce the number of categories by creating Var4 since it seems that anyone who does not have veggies it more likely to suffer from heart disease or attack. With reagrds to Var5 I found it interesting that someone with high alcohol consumption is less likely to have heart disease or attack comnpare to someone who doesn't smoke or not had heart disease, this could just mean people who have high alcohol consumption will have liver issues before having heart disease or attack. Interstingly someone who is underwigeight is more likely to have heart disease or attack than someone who is classed as overweight.

```{r echo=FALSE}
plots<-plot_grid(TrainPlot1MentHlth2 + theme(legend.position = "none"),  TrainPlot1PhysHlth2 + theme(legend.position = "none") + ylab(""),  TrainPlot1Education2 + theme(legend.position = "none") , TrainPlot1Age2 + theme(legend.position = "none") +  ylab(""))
legend <- get_legend(TrainPlot1Education)
plot_grid(plots, legend, rel_widths = c(2, 0.2))
```

With regards to mental health is seems as that having some mental health (between 1 and 9) has a positive effect on heart disease or attack. Not surprisingly we saw that people who exercise more frequently are less likely to have heart disease or attack. We saw that a person with lower education is more likely to have heart disease or attack however this could be linked to income and standard or living (people with high income are less likely). Not surprisingly the older you are the more likely you are to have heart disease or attack since the person has had more time to have heart disease or attack.

### Model creation

In short form I decided to create 3 classification models using to different algorithm (Random Forest, GLM Logistic regression using forward selction based on AIC and CART) using the Train2 dataset. Each modek was assessed in the Train and Test dataset and based on the best model create out of the 3 (GLM) I then proceeded to see if I could improve the model further by creating dummy variables (converting a single cateogorical variable into multiple variables with a binary response) of the variables selected in my GLM model and again running the forward selection as my 2nd model. Then I created a 3rd GLM model that uses backward selecion based on my 2nd model. I then created a 4th GLM model that uses forward and backwards selection of the variables used in my 2nd model. A 5th GLM model created by removing any insignificant variables (so that all variables incuded in the model have a p-value of \<0.05) from the 3rd model and then a 6th GLM model was created which is the same as the 4th GLM model but all insignifcant variables. I then compared the results from the GLM models based on the Train and Test datasets I concluded that that the 5th GLM model is the best since there was minimal differences between the results (between accuracy, sensitivity, precision and Somers D ) but it had the lowest number of variables. I then proceed to adjuist my cuttoff value in the 5th GLM model that determines to identify people with heart disease or attack, the best cuttoff value with determined by the highest Somers D statistic. I then assessed my optimised 5th GLM model against my holdout sample which produced similar results compared to the Train and Test datasets.

The metrics I used to determine how each model perfomed on the Train and Test datasets:

-   NVars - This is the number of variables used in the model, the lower the better.

-   Accuracy - Shows the proportion of correctly classified outcomes, the higher the better but because I have an imbalanced dataset with regards to heart disease or attack it's acceptable to reduce this figure if I'm identifying more people with heart disease or attack.

-   Sensitivity - This is the true positives divided by the true positives plus the false negatives, the higher the better.

-   Precision - This is the true positives divided by the true positives plus the false positives, the higher the better.

-   NcountTP - The number of true positives identified in the model i.e. the higher the better.

-   Somers D - This is a measure of the strength and direction of the association between an ordinal dependent variable and an ordinal independent variable [4]. the score ranges between -1 and 1 i.e. the higher the better and holds the most weight with regards to picking the best model.

### Results

##### Here is the table of my summary results based on the 3 initial models I created (Random Forest, GLM & CART) based on the train and test datasets:

```{r echo=FALSE}
ModelResultsSummary  %>% filter(row_number() <= n()-6)
```

From the above table you will see that the GLM model performed the best when it came to the Somers D statistic plus it also identified the most true positives i.e. I explored GLM logistic regression further.

As stated in the first paragraph of this section I created another 3 GLM models based on the variables from my inintal model except that I had used various selection methods (forward, backwards and both directions). I found that there was no change to the model between the forward and backward models. I could see that there are insignificant variables i.e. I created another 2 GLM models that removed the insignificant variables.

##### Here is the table of my summary results based on the forward (GLM2), backward (GLM3), both direction (GLM4), removed insigniifcant variables from GLM3 (GLM5) and removed insiginficant variables from GLM4 (GLM6) models based on the train and test datasets:

```{r echo=FALSE}
ModelResultsSummary[c(4:8),]
```

From the above table you will see that the results are very similar to each except that the GLM5 model used fewer (removed insignificant variables from GLM3) number of variables i.e. I believe this is the best model I have and have decided this to be my final model. Look below to the see the model summary of my GLM5:

```{r echo=FALSE}
summary(fit.glm5)
```

Here is a graphical view (Train - LHS and Test - RHS) below of the performance of the Train and Test data, essentially I have scored all records in the dataset, I then put the probability scores into 10 bins (1 to 10) and then added the proportion of people that have been identified as having heart disease or attack. You will see the higher the probability score is the better the model perfomed.

```{r echo=FALSE}
plot_grid(
ggplot()+
    geom_bar(data=GLM5SummaryResultsTrain, aes(x=PredCatGLM5, y=Freq), stat="identity", position="identity", show.legend=FALSE)+
    geom_line(data=GLM5SummaryResultsTrain, aes(x=PredCatGLM5, y=Target*2*max(Freq)), size=1.5, color="red") 
,
ggplot()+
    geom_bar(data=GLM5SummaryResultsTest, aes(x=PredCatGLM5, y=Freq), stat="identity", position="identity", show.legend=FALSE)+
    geom_line(data=GLM5SummaryResultsTest, aes(x=PredCatGLM5, y=Target*2*max(Freq)), size=1.5, color="red") 
)
```

I proceeded to fine tune my cuttoff value (what the probability score needs to be) that determines whether a person should be classified as having heart disease or attack. To do this I ran a sequence between 0.01 to 1 in increments of 0.01 which I effectively changed the cuttoff value and then I produced a list of statistics. I based my cuttoff value on the highest Somers D statistic on the Train & Test dataset where I took the average best cuttoff values. Here is a graphical view.

```{r echo=FALSE}
par(mfrow = c(1, 2))
plot(cutoffsTrain$Dxy)
plot(cutoffsTest$Dxy)
```

```{r echo=FALSE}
cutoffsTrain[which.max(cutoffsTrain$Dxy),]
```

```{r echo=FALSE}
cutoffsTest[which.max(cutoffsTest$Dxy),]
```

I then updated my model summary table and will see that my optimised 5th GLM model has perfomed best with regards to Somers D and identifying the most true positives.

```{r echo=FALSE}
ModelResultsSummary
```

I then ran my 5th GLM model using my cuttoff value against my holdout dataset and got similar results compared to my Train and Test data.

```{r echo=FALSE}
#model summary results Holdout
cm.glm5_HoldoutOptimised
#sommers d 
somers2(Holdout_Results$ResultHDA1GLM5,  Holdout$HeartDiseaseorAttack)
```

# Conclusion

My final model has a moderate Somers D statistic value and was able to capture the majority of people who have heart disease or attack. At the same time I also predicted many people that would have it but haven't had it however I have identified these people most at risk of having heart disease or attack and suggest that these people are examined by a medical professional to determine the risk they are at of having heart disease or attack.

If I were to do the project again I would look to create multiple models for based on the age group a person sits in as we know that the older a person is the more likely they are to have heart disease or attack and the at the same time nobody has control over this factor.

# References

1)  I got my dataset from kaggle from the following URL: <https://www.kaggle.com/datasets/alexteboul/heart-disease-health-indicators-dataset>

2)  To download the Kaggle dataset as a csv file, use the link below: <https://www.kaggle.com/datasets/alexteboul/heart-disease-health-indicators-dataset/download?datasetVersionNumber=3>

3)  BMI categories coming from the NHS UK which have been used to create the BMI2 variable: <https://www.nhs.uk/conditions/obesity/>

4)  What is Somers' D:

    <https://www.statology.org/somers-d/>
